### 2008-11-03 papers/jss/dmm4.tex
added papers directory (lots of fighting with svn ...)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@234 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%&LaTeX

\documentclass[a4paper,12pt,doc]{apaIngmar} % nobf, doc, apa, man

\usepackage[]{amsmath, amsfonts, amstext, amsthm}

\usepackage{amssymb}

\usepackage[]{graphics}

\usepackage{graphicx}

\usepackage{epsfig}

\usepackage{epstopdf}

\newcommand{\citep}{\cite}

\newcommand{\citet}{\citeA}

\newcommand{\mat}{\mathbf}

\newcommand{\vc}{\mathbf}

\title{DMM4: Dependent Mixture Models in S4 Classes}

\author{Ingmar Visser}
\date{\today}

\affiliation{Department of Psychology, University of Amsterdam\\

 Correspondence concerning this article should be addressed to:
 \\
 Ingmar Visser \\
 Department of Psychology, University of Amsterdam \\
 Roetersstraat 15 \\
 1018 WB Amsterdam \\
 phone: +31 (20) 5256735 \\
 fax: +31 (20) 6390279 \\
 email: i.visser@uva.nl}

\renewcommand{\baselinestretch}{1.5}

\shorttitle{DMM4: Dependent Mixture Models in S4}

\abstract{DMM4 implements mixture models and latent Markov models for
mixed indicator from various distributions.  Optimization can be
subjected to general linear constraints and is done using nlm, optim
or NPSOL. Covariates can be regressed on model parameters through
identity or logit link functions.}

\begin{document}

\bibliographystyle{apacite}

\maketitle

\newpage

In this paper an outline is provided for a program that computes
parameter estimates for the latent/hidden Markov mode (LMM) and for
latent transition analysis (LTA), both univariate and multivariate.
The difference between the two types of model is not a principled one
but rather of degree, being N is small and T large in LMM and vice
versa in LTA. Even so, when T is indeed large as is usually the case
in the hidden Markov model literature, the transition matrix is taken
to be stationary.  The basis for the current approach is to use the
hidden Markov model (HMM) estimation equations since these are
suitable for long time series which is not the case for the standard
EM algorithm used for latent class analysis (LCA) and LTA which soon
runs into the problem of sparse contingency tables and underflow
problems during computation.  The approach is illustrated using a
number of simulated and real data examples.

\section{Introduction}

The model that is considered in this paper is the multivariate (or
multi indicator) latent Markov model.  The main contribution of this
paper is the extension of estimation algorithms to allow for arbitrary
length time series of observations.  In traditional estimation methods
this is impossible due to underflow problems.  Moreover, the
estimation procedure that I present is not based on the complete
contingency table, which is impossible or impractical in the case of
long time series.  In particular, one might want to model a single
time series.  The usual approach to latent Markov models is then faced
with a huge contingency table, $2^{T}$ cells for a binary variable
measured at $T$ time points, in which only a single cell has a one and
the others are zero.  Howzthat for sparsity?

The model can be seen as an extension of several other models, in
particular, the hidden Markov model (HMM), the latent class model
(LCM), the latent transition model (LTM) and also the simple Markov
model (MM).  Markov models (and their subsequent refinements) have a
long history in psychology, mainly in the psychology of learning
\citep{Wickens1982}.  These models have in common that they are all
used to describe one or more sequences of responses $O_{1}, \ldots,
O_{T}$, where each of the $O_{t}$'s may be multivariate.  In this
paper mixed categorical and continuous observations are considered.

The model concerned in this paper is the combination of the latent
class model and the hidden Markov model.  There is quite some work
already on this model, but it is usually limited to small numbers of
repeated meausurements, in which case this model is usually called the
latent transition model: it is the repeated measurements extension of
the latent class model.  It is customary in those models to have
time-varying (transition) parameters.  In the time series literature
on the other hand, where hidden Markov models are frequently used,
transition parameters are taken to be time-homogeneous.  This
assumption is necessary when analyzing single time-series since at
each time $t$ only a single transition is observed.  The multi-variate
latent Markov model is defined by the following elements:
\begin{enumerate}
	\item a set $\vc{S}$ of latent states $S_{i},\, i=1, \ldots , n$,

	\item a matrix $\mat{A}$ of transition probabilities $a_{ij}$ for
	the transition from state $S_{i}$ to state $S_{j}$,

	\item a set $\vc{B}$ of observation functions $b_{j}(\cdot)$ that
	provide the conditional probabilities associated with latent state
	$S_{j}$,

	\item a vector $\pmb{\pi}$ of latent state initial probabilities
	$\pi_{i}$.
\end{enumerate}
When transitions are added to the latent class model, it is more
appropriate to refer to the classes as states.  The word class is
rather more associated with a stable trait-like attribute whereas a
state can change over time.

The data has the general form $O_{1}^{1}, \ldots, O_{1}^{m},
O_{2}^{1}, \ldots, O_{2}^{m}, \ldots, O_{T}^{1}, \ldots, O_{T}^{m}$
for an $m$-variate time series of length $T$ and independent
realizations of such series.  It is important to note that the latent
Markov model is commonly associated with data of this type, but the
estimation procedures that are used are not suitable for long time
series due to underflow problems.  Conversely the hidden Markov model
is only used for univariate time series and multiple dependent time
series or multivariate time series have not been considered in
previous literature on hidden Markov models.  In the next section, the
likelihood and estimation procedure for this model is described, given
data of the above form, for arbitrary $m$ and $T$.

\section{Computing likelihood, gradients and Hessian}

Single-indicator latent class models.  Estimation logic is different:
the starting point is not a complete contingency table, which is then
weeded.  The starting point is a stochastic automaton which produces
time series.  In time series analysis, one does not expect a complete
contingency table; in the contingency table logic, a single time
series produces an enormous table with only a single filled cell.

For computing the likelihood and its partial derivatives, I follow the
strategy used by \cite{Lystig2002}.  They developed a method to
compute exact gradients and information matrix by reformulating the
forward procedure of the Baum-Welch algorithm for estimating hidden
Markov model parameters.  Scores and information are computed at a
single pass through the data, using the intermediate variables from
the forward recursion.  Their article only discusses univariate time
series.  Our aim in this section is to generalize their results to
multiple dependent and independent time series.  With the gradients
available, it is possible to use general purpose optimization routines
that allow box constraints and arbitrary (non-) linear constraints.

\subsection{Likelihood}

The loglikelihood of hidden Markov models is usually computed by the
so-called forward-backward algorithm \citep{Baum1966,Rabiner1989}, or
rather by the forward part of this algorithm.  \cite{Lystig2002}
change the forward algorithm in such a way as to allow computing the
gradients of the loglikelihood at the same time.  They start by
rewriting the likelihood as follows (for ease of exposition the
dependence on the model parameters is dropped):
\begin{equation}
	L_{T} = Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) = \prod_{t=1}^{T}
	Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, \vc{O}_{t-1}),
	\label{condLike}
\end{equation}
where $Pr(\vc{O}_{1}|\vc{O}_{0}):=Pr(\vc{O}_{1})$.  Note that for a
simple Markov chain these probabilities reduce to
$Pr(\vc{O}_{t}|\vc{O}_{1},\ldots, \vc{O}_{t-1}) =
Pr(\vc{O}_{t}|\vc{O}_{t-1})$.  The log-likelihood can now be expressed
as:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log[Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots,
	\vc{O}_{t-1})].
	\label{eq:condLogl}
\end{equation}

To compute the log-likelihood, \cite{Lystig2002} define the following 
(forward) recursion:
\begin{align}
	\phi_{1}(j) &:= Pr(\vc{O}_{1}, S_{1}=j) = \pi_{j}
	b_{j}(\vc{O}_{1}) \label{eq:fwd1} \\
\begin{split}
	\phi_{t}(j) &:= Pr(\vc{O}_{t}, S_{t}=j|\vc{O}_{1}, \ldots,
	\vc{O}_{t-1}) \\
	&= \sum_{i=1}^{N} [\phi_{t-1}(i)a_{ij}b_{j}(\vc{O}_{t})] \times
	(\Phi_{t-1})^{-1},
	\label{eq:fwdt} 
\end{split} 
\end{align}
where $\Phi_{t}=\sum_{i=1}^{N} \phi_{t}(i)$.  Combining
$\Phi_{t}=Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, \vc{O}_{t-1})$, and
equation~(\ref{eq:condLogl}) gives the following expression for the
log-likelihood:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log \Phi_{t}.
\end{equation}

The above forward recursion can readily be generalized to mixture
models, in which it is assumed that the data are realizations of a
number of different LMMs and the goal is to assign posterior
probabilities to sequences of observations.  This situation occurs,
for example, in learning data where different learning strategies may
lead to different answer patterns.  From an observed sequence of
responses, it may not be immediately clear from which learning process
they stem.  Hence, it is interesting to consider a mixture of latent
Markov models which incorporate restrictions that are consistent with
each of the learning strategies.

\subsection{Likelihood for mixtures}

The easiest way to compute the likelihood of a mixture of LMMs is by
realizing that such a mixture is itself an LMM with a number of
constraints on the initial parameters and the transition matrix.  For
example, consider a mixture of two LMMs, in which one component has
two states and the other three states.  Later in this paper I will
present an example of such a mixture when analyzing discrimination
learning data.  The assumption is that children and adults use
different strategies in such a task which can be translated into
different latent Markov models.  In this example the first component
has two states and the following parameters:
\begin{gather*}
\mat{A} =\begin{pmatrix} 
1 & 0\\ \alpha & 1-\alpha
\end{pmatrix} \quad \text{and} \quad
\pmb{\pi}=\begin{pmatrix} \pi_{1}\\ \pi_{2} \end{pmatrix}.
\end{gather*}
The response parameters $b$ are all to freely estimated. The second 
component has three  states and the following parameters:
\begin{gather*}
\mat{A} =\begin{pmatrix} 
1 & 0 & 0\\ 0 & g & 1-g \\ \alpha & \beta & \gamma
\end{pmatrix} \quad \text{and} \quad
\pmb{\pi}=\begin{pmatrix} \pi_{1}\\\pi_{2}\\\pi_{3}  \end{pmatrix}.
\end{gather*}
By combing the transition matrices and the initial vectors the 
following model results:
\begin{gather*}
\mat{A} =\begin{pmatrix} 
1 & 0 & 0 & 0 & 0\\
\alpha & 1-\alpha & 0 & 0 & 0 \\
0 & 0 & 1 & 0 & 0
\\0 & 0 & 0 & g & 1-g 
\\ 0 & 0 & \alpha & \beta & \gamma
\end{pmatrix} \quad \text{and} \quad
\pmb{\pi}=\begin{pmatrix} \pi_{1}^{1}\\ \pi_{2}^{1}  \\ 
\pi_{1}^{2}\\\pi_{2}^{2}\\\pi_{3}^{2} \end{pmatrix},
\end{gather*}
whith the constraint that the $\pi$'s sum to unity.  It can easily be
seen that computing the log-likelihood of such a model leads to many
unneccessary computations because the transition matrix $\mat{A}$
contains blocks of zeroes that have to be multiplied $T$ times for
each time series under consideration.  Therefor, below the
log-likelihood is computed as a real mixture of the components above
with mixture parameters $p_{k}$.

To compute the likelihood of a mixture of $K$ models, define the
forward recursion variables as follows (these variables now have an
extra index $k$ indicating that observation and transition
probabilities are from latent model $k$):
\begin{align}
\begin{split}
\phi_{1}(j_{k}) &=  Pr(\vc{O}_{1}, 
S_{1}=j_{k})=p_{k}\pi_{j_{k}}b_{j_{k}}(\vc{O}_{1}).
\end{split}\label{eq:fwd1mix} \\
\begin{split}
\phi_{t}(j_{k})   &=   Pr(\vc{O}_{t}, S_{t}=j_{k}|\vc{O}_{1}, \ldots, 
\vc{O}_{t-1}) \\
			&= \left[ \sum_{k=1}^{K} \sum_{i=1}^{n_{k}} \phi_{t-1}(i_{k}) 
			a_{ij_{k}}b_{j_{k}}(\vc{O}_{t}) \right] \times (\Phi_{t-1})^{-1},
\end{split}\label{eq:fwdtmix} 
\end{align}
where $\Phi_{t} =  \sum_{k=1}^{K}\sum_{i=1}^{n_{k}} 
\phi_{t}(j_{k})$.  Note that the double sum over $k$ and $n_{k}$ is 
simply an enumeration of all the states of the model. Now, because 
$a_{ij_{k}}=0$ whenever $S_{i}$ is not part of component $k$, the sum 
over $k$ can be dropped and hence equation~\ref{eq:fwdtmix} reduces 
to:
\begin{equation}
	\phi_{t}(j_{k}) = \left[ \sum_{i=1}^{n_{k}} \phi_{t-1}(i_{k}) 
			a_{ij_{k}}b_{j_{k}}(\vc{O}_{t}) \right] \times (\Phi_{t-1})^{-1}
\end{equation}

Writing this out in matrices and vectors and rearranging provides 
useful expressions for implementational  purposes, for each $k$ we 
get:
\begin{align}
	\vc{\phi}_{1} &=  p_{k} \times \pmb{\pi} \otimes 
\vc{b}(\vc{O}_{1}),\\
	\vc{\phi}_{t} &= \left[ \mat{A}^{t} \times \pmb{\phi}_{t-1} \right ] 
		\otimes \vc{b} (\vc{O}_{t}) \times (\Phi_{t-1})^{-1},
\end{align}
where $^{t}$ denotes matrix tranpose, $\times$ denotes matrix/vector 
multiplication and $\otimes$ denotes the Hadamard product of 
vectors/matrices.

\paragraph{Computational considerations} 
From equations~(\ref{eq:fwd1}--\ref{eq:fwdt}), it can be seen that
computing the forward variables, and hence the log-likelihood, takes
$O(Tn^{2})$ computations, for an $n$-state model and a time series of
length $T$.  Consider the mixture of two components, one with two
states and the other with three states, as in the example above.
Using equations~(\ref{eq:fwd1}--\ref{eq:fwdt}) to compute the
log-likelihood of this model one needs $O(Tn^{2})=O(T\times 25)$
computations whereas with the mixture
equationsequations~(\ref{eq:fwd1mix}--\ref{eq:fwdtmix}), one only
needs $\sum_{n_{i}} O(n_{i}^{2}T)$ computations, in this case $O(T
\times 13)$.  So, it can be seen that in this easy example the
computational cost is almost halved.

Note that in above equations $b_{j}(\vc{O}_{t})$ factors into
$\prod_{i=1}^{m} b_{j} (O_{t}^{i})$, as is the case in the latent
class model.

%\subsection{Posterior probabilities}

%Viterbi algorithm?

% %Using the equations above for the likelihood of single and mixture 
% models, arriving at the posterior probabilities of membership of the 
% components can be easily accomplished.

% %The posterior probability is the probability that data $\vc{O}$ are 
% generated by mixture component $k$, i.e., $Pr(C=k|\vc{O})$, where $C$ 
% denotes the mixture component. This probability can be expressed as:
%\begin{eqnarray}
%	Pr(C=k|\vc{O}) &=& \frac{Pr(C=k) Pr(\vc{O}|C=k)}
%	{\sum_{k=1}^{K} Pr(C=k)Pr(\vc{O}|C=k)} \\
%	&=& \frac {p_{k} l_{TMC=k}}
%	{\sum}
%\end{eqnarray}

\subsection{Gradients}

\newcommand{\fpp}{\frac{\partial} {\partial \lambda_{1}}}

See equations 10--12 in \cite{Lystig2002} for the score recursion
functions of the hidden Markov model for a univariate time series.
Here the corresponding score recursion for the multivariate mixture
case are provided.  The $t=1$ components of this score recursion are
defined as (for an arbitrary parameter $\lambda_{1}$):
\begin{align}
\psi_{1}(j_{k};\lambda_{1}) &:=  \fpp Pr(\vc{O}_{1}|S_{1}=j_{k}) \\
\begin{split} 
	&= \left[  \fpp p_{k} \right] \pi_{j_{k}}b_{j_{k}}(\vc{O}_{1}) + 
	p_{k}\left[ \fpp \pi_{j_{k}} \right] b_{j_{k}}(\vc{O}_{1}) \\
	& \qquad  + p_{k}\pi_{j_{k}} \left[ \fpp 
b_{j_{k}}(\vc{O}_{1})\right],
\end{split} \label{eq:psi1}
\end{align}
and for $t>1$ the definition is:
\begin{align}
	\psi_{t}(j_{k};\lambda_{1}) & = \frac{\fpp Pr(\vc{O}_{1}, \ldots,
	\vc{O}_{t}, S_{t}=j_{k})} {Pr(\vc{O}_{1}, \ldots, \vc{O}_{t-1})}
	\\
	\begin{split} 
		& = \sum_{i=1}^{n_{k}} \Bigg\{
		\psi_{t-1}(i;\lambda_{1})a_{ij_{k}} b_{j_{k}}(\vc{O}_{t}) \\
			&\qquad +\phi_{t-1}(i) \left[ \fpp a_{ij_{k}} \right]
			b_{j_{k}} (\vc{O}_{t}) \\
			&\qquad +\phi_{t-1}(i)a_{ij_{k}}  \left[ \fpp b_{j_{k}} 
			(\vc{O}_{t}) \right] \Bigg\} 
			\times (\Phi_{t-1})^{-1}.
		\end{split} 
	\label{eq:psit}
\end{align}

Writing this out in matrix notation (and dropping the $k$-subscripts 
for components) we get for $t=1$ and for each $k$:
\begin{eqnarray}
	\pmb{\psi}_{1}(\lambda_{1}) = \vc{\fpp} p_{k} \times \pmb{\pi} 
	\otimes \vc{b}(O_{1})+ p_{k} \times \vc{\fpp}\pmb{\pi} \otimes
	\vc{b}(O_{1}) + p_{k} \times \pmb{\pi} \otimes \fpp \vc{b}(O_{1}),
\end{eqnarray}
and for $\pmb{\psi}_{t}$ we get: 
\begin{equation}
	\begin{split} 
		\pmb{\psi}_{t}(\lambda_{1}) &= \left[ \mat{A}^{t} \times
		\pmb{\psi_{t-1}}(\lambda_{1}) \right ] \otimes \vc{b}(O_{t}) +
		\left [ \left[ \fpp \mat{A} \right ]^{t} \times
		\pmb{\phi}_{t-1} \right ] \otimes \vc{b}(O_{t}) \\
		& \qquad + \left [ \mat{A}^{t} \times \pmb{\phi}_{t-1} \right]
		\otimes \fpp \vc{b}(O_{t}).
	\end{split}
\end{equation}

Using above equations, \cite{Lystig2002} derive the following equation 
for the partial derivative of the likelihood:
\begin{equation}
	\fpp l_{T}=
	\frac{\mathbf{\Psi}_{T}(\lambda_{1})}{\mathbf{\Phi}_{T}},
\end{equation}
where $\Psi_{t}=\sum_{k=1}^{K} \sum_{i=1}^{n_{k}} 
\psi_{t}(j_{k};\lambda_{1})$. 
Starting from the equation from the logarithm of the likelihood, this 
is easily seen to be correct: 
\begin{eqnarray}
	\fpp \log Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) &=& Pr(\vc{O}_{1},
	\ldots, \vc{O}_{T})^{-1} \fpp Pr(\vc{O}_{1}, \ldots, \vc{O}_{T})
	\\
	&=& Pr(\vc{O}_{1}, \ldots, \vc{O}_{T})^{-1} \Psi_{T}(\lambda_{1})
	Pr(\vc{O}_{1}, \ldots, \vc{O}_{T-1}) \\
	&=& \frac{Pr(\vc{O}_{1}, \ldots, \vc{O}_{T-1})}{Pr(\vc{O}_{1},
	\ldots, \vc{O}_{T})} \Psi_{T} (\lambda_{1}) \\
	&=&  \frac{\mathbf{\Psi}_{T}(\lambda_{1})}{\mathbf{\Phi}_{T}}.
\end{eqnarray}

Having provided the global computations above we now need expressions
for $\fpp p_{k}$, $\fpp \pi_{i}$, $\fpp a_{ij}$, and $\fpp
b_{i}(\vc{O}_{t})$, which are dependent on the particular
$\lambda_{1}$ under consideration.  This generates special cases for
each combination of $\lambda_{1}$ and each of the components of the
gradient computation.  These cases are worked out below for each of
the choices $\lambda_{1} \in \pmb{\lambda}=\{
\vc{p},\pmb{\pi},\mat{A},\vc{b} \}$.

%%p-k cases
\paragraph{Case 1: $\fpp p_{k}$}

\begin{align} 
        \fpp p_{k} = \left\{ \begin{array}{ll} 
            1 & \mbox{if $\lambda_{1}=p_{k}$} \\ 
            0 & \mbox{otherwise} 
            \end{array} \right. 
\end{align}

%%pipi-k cases

\paragraph{Case 2: $\fpp \pi_{i}$ }

\begin{align} 
        \fpp \pi_{i}^{k} = \left\{ \begin{array}{ll} 
            1 & \mbox{if $\lambda_{1}=\pi_{i}^{k}$} \\ 
            0 & \mbox{otherwise} 
            \end{array} \right. 
\end{align}

\paragraph{Case 3: $\fpp a_{ij}$ }

\begin{align} 
        \fpp a_{ij}^{k}= \left\{ \begin{array}{ll} 
            1 & \mbox{if $\lambda_{1}=a_{ij}^{k}$} \\ 
            0 & \mbox{otherwise} 
            \end{array} \right. 
\end{align}

The cases below for $\fpp b_{j}(\vc{O}_{t})$ are slightly more
interesting.  First, remember that in the case of an $m$-variate time
series of observations, $\fpp b_{j}(\vc{O}_{t})=\fpp b_{j}(O_{t}^{1},
\ldots, O_{t}^{m})$, of which we need partial derivatives for
arbitrary parameters $\lambda_{1}$.  First note that because of local
independence we can write:
\begin{equation}
	\fpp [b_{j}(O_{t}^{1}, \ldots, O_{t}^{m})]= \frac{\partial} 
{\partial \lambda_{1} } [b_{j}(O_{t}^{1})] \times  
[b_{j}(O_{t}^{2})], \ldots,  [b_{j}(O_{t}^{m})].  
\end{equation}
Applying the chain rule for products we get:
\begin{equation}
	\fpp [b_{j}(O_{t}^{1}, \ldots, O_{t}^{m})] =
	\sum_{l=1}^{m} \left[ \prod_{i=1, \ldots, \hat{l}, \ldots, m} 
b_{j}(O_{t}^{i}) \right] \times
	\fpp  [b_{j}(O_{t}^{l})],
	\label{partialProd}
\end{equation}
where $\hat{l}$ means that that term is left out of the product.
These latter terms, $\frac{\partial} {\partial \lambda_{1} }
[b_{j}(O_{t}^{k})]$, are easy to compute given either multinomial or
gaussian observation densities.  As a consequence, this provides
gradients for the multivariate time series model for mixed categorical
and gaussian observations, e.g.\ correct/incorrect responses and
reaction times, and a categorical latent state or class.  Hence, we
need the partial derivatives $\fpp b_{j}({O}_{t}^{l})$ instead of
$\fpp b_{j}(\vc{O}_{t})$ (if $\lambda_{1} \in \{ \vc{p},
\pmb{\pi},\mat{A} \} $, then $\fpp b_{j}=0$).

\paragraph{Case 4: $\fpp b_{j}({O}_{t}^{l})$ and $\lambda_{1} \in
\vc{b}$.} These derivatives depend on the response function of the
item, i.e.\ whether the item is multinomial or gaussian (other
itemtypes are easy to add).  For the derivatives of the response
functions there are two cases depending on the type of the response
function.  For a gaussian item we have the following:
\begin{align} 
	\fpp b_{j_{k}}(O_{t}^{l})= \left\{
	\begin{array}{ll} 
		b_{j_{k}}(O_{t}^{l}) \times \left[
		\frac{O_{t}-\mu}{\sigma^{2}} \right] & \mbox{if $\lambda_{1} =
		\mu_{j_{k}}^{l} $ } 
		\\
		b_{j_{k}}(O_{t}^{l}) \times \left[
		\frac{(O_{t}-\mu)^{2}}{\sigma^{3}} - \frac{1}{\sigma} \right]
		& \mbox{if $\lambda_{1} = \sigma_{j_{k}}^{l} $} 
		\\
		0 & \mbox{otherwise} \end{array} \right.
\end{align}

For an $c$-category item $l$ with response probabilities $p_{1}, 
\ldots, p_{c}$ we have the following:
\begin{align} 
	\fpp b_{j_{k}}(O_{t}^{l})= \left\{ 
	\begin{array}{ll} 
		1 & \mbox{if $\lambda_{1} = p_{i}$ \& $O_{t}=i$}  \\ 
		0 & \mbox{otherwise} 
	\end{array} 
	\right\}
\end{align}

%\newpage
%\section*{Author note}
%We would like to thank many people.

\bibliography{all}

\end{document}