### 2008-11-03 papers/vignette/depmix-intro.tex

\documentclass[a4paper]{article}

\usepackage[]{amsmath, amsfonts, amstext, amsthm}

\usepackage{amssymb}

\usepackage[]{graphics}

\usepackage{graphicx}

\usepackage{epsfig}

\usepackage{epstopdf}

%\batchmode

\usepackage[round]{natbib}

\title{depmixS4: An R-package for fitting mixture models on mixed
multivariate data with Markov dependencies}

\author{Ingmar Visser \& Maarten Speekenbrink\thanks{
Correspondence concerning this manual should be adressed to:
Ingmar Visser, Department of Psychology, University of Amsterdam,
Roetersstraat 15, 1018 WB, Amsterdam, The Netherlands}\\
Department of Psychology, University of Amsterdam\\
i.visser@uva.nl \\ 
Department of Psychology, University College London 
\\ m.speekenbrink@ucl.ac.uk}

\date{\today}

\newcommand{\vc}{\mathbf}
\newcommand{\mat}{\mathbf}

 \newcommand{\pkg}{\texttt}
 \newcommand{\code}{\texttt}

\begin{document}

\bibliographystyle{plainnat}

\maketitle

\begin{abstract}

	\pkg{depmixS4} implements a general framework for definining and
	fitting hidden Markov mixture model in the R programming language
	\citep{R2007}.  This includes standard Markov models,
	latent/hidden Markov models, and latent class and finite mixture
	distribution models.  The models can be fitted on mixed
	multivariate data with multinomial and/or gaussian distributions.
	Parameters can be estimated subject to general linear constraints.
	Parameter estimation is done through an EM algorithm or by a
	direct optimization approach with gradients using the
	\code{Rdonlp2} optimization routine when contraints are imposed on
	the parameters.  A number of illustrative examples are included.

\end{abstract}

\newpage
\tableofcontents
\newpage

\section{Introduction}

Markov and latent Markov models are frequently used in the social
sciences, in different areas and applications.  In psychology, they
are used for modelling learning processes, see \citet{Wickens1982},
for an overview, and \citet{Schmittmann2006} for a recent application.
In economics, latent Markov models are commonly used as regime
switching models, see e.g.\ \citet{Kim1994} and \citet{Ghysels1994}.
Further applications include speech recognition \citep{Rabiner1989},
EEG analysis \citep{Rainer2000}, and genetics \citep{Krogh1998}.  In
those latter areas of application, latent Markov models are usually
referred to as hidden Markov models.

The \pkg{depmixS4} package was motivated by the fact that Markov models
are used commonly in the social sciences, but no comprehensive package
was available for fitting such models.  Common programs for Markovian
models include Panmark \citep{Pol1996}, and for latent class models
Latent Gold \citep{Vermunt2003}.  Those programs are lacking a number
of important features, besides not being freely available.  There are
currently some packages in R that handle hidden Markov models but they
lack a number of features that we needed in our research.  In
particular, \pkg{depmixS4} was designed to meet the following goals: 
\begin{enumerate}

	\item to be able to handle parameter estimates subject to general
	linear (in)equality constraints

	\item to be able to fit transition models with covariates, i.e.,
	to have time-dependent transition matrices

	\item to be able to include covariates in the prior or initial
	state probabilities of models

	\item to allow for easy extensibility, in particular, to be able
	to add new response distributions, both univariate and
	multivariate, and similarly to be able to allow for the addition
	of other transition models, e.g., continuous time observation
	models

\end{enumerate}

Although \pkg{depmixS4} is designed to deal with
longitudinal or time series data, for say $T>100$, it can also handle
the limit case with $T=1$.  In those cases, there are no time
dependencies between observed data, and the model reduces to a finite
mixture model, or a latent class model.  Although there are other
specialized packages to deal with mixture data, one specific feature
that we needed ourselves which is to the best of our knowledge not
available in other packages is the possibility to include covariates
on the prior probabilities of class membership.  In the next section,
an outline is provided of the model and the likelihood equations.

\subsection*{Acknowledgements}

Ingmar Visser was supported by an EC Framework 6 grant, project 516542
(NEST).  Maarten Speekenbrink was supported by the ESRC Centre for
Economic Learning and Social Evolution (ELSE).  Han van der Maas
provided the speed-accuracy data \cite{Maas2005b} and thereby
neccessitated implementing models with time-dependent covariates.
Brenda Jansen provided the balance scale data set \citep{Jansen2001}
which was the perfect opportunity to test the covariates on the prior
model parameters.  The examples in the help files use both of these
data sets.

% Maartje Raijmakers \cite{Raijmakers2001} provided
% the discrimination data set.

\section{Dependent mixture models}

%Only later did I find out that \citet{Leroux1992} already
%coined the term dependent mixture models in an application with hidden
%Markov mixtures of Poisson count data.

The data considered here, has the general form $O_{1}^{1}, \ldots,
O_{1}^{m}$, $O_{2}^{1}, \ldots, O_{2}^{m}$, \ldots, $O_{T}^{1}, \ldots,
O_{T}^{m}$ for an $m$-variate time series of length $T$.  As an
example, consider a time series of responses generated by a single
subject in a reaction time experiment.  The data consists of three
variables, reaction time, accuracy and a covariate which is a pay-off
factor which determines the reward for speed and accuracy.  These
variables are measured on 168, 134 and 137 occasions respectively 
(in Figure~\ref{fig:speed} the first part of this series is plotted).

\begin{figure}[htbp]
  \begin{center}
	  \scalebox{0.6}{\includegraphics*{speed1.pdf}}
	  \caption{Reaction times, accuracy and pay-off values for
	  the first series of responses in dataset \texttt{speed}.}
	  \label{fig:speed}
  \end{center}
\end{figure}

The latent Markov model is commonly associated with data of this type,
albeit usually only multinomial variables are considered.  However,
common estimation procedures, such as those implemented in
\citet{Pol1996} are not suitable for long time series due to underflow
problems.  In contrast, the hidden Markov model is typically only used
for `long' univariate time series.  In the next section, the
likelihood and estimation procedure for the hidden Markov model is
described, given data of the above form.  These models are called
dependent mixture models because one of the authors (Ingmar Visser)
thought it was time for a new name for these models.

The dependent mixture model is defined by the following elements:
\begin{enumerate}

	\item a set $\vc{S}$ of latent classes or states $S_{i},\, i=1,
	\ldots , n$,

	\item matrices $\mat{A}_t$ of transition probabilities $a_{ij,t}$ for
	the transition from state $S_{i}$ to state $S_{j}$ at time $t$,

	\item a set $\vc{B}_t$ of observation functions $b_j^k(\cdot)$ that
	provide the conditional probabilities of observations $O_{t}^k$ 
	associated with latent state $S_{j}$,

	\item a vector $\pmb{\pi}$ of latent state initial probabilities
	$\pi_{i}$
\end{enumerate}
When transitions are added to the latent class model, it is more 
appropriate to refer to the classes as states. The word class is 
rather more associated with a stable trait-like attribute whereas a 
state can change over time.

\subsection{Likelihood}

The log-likelihood of hidden Markov models is usually computed by the
so-called forward-backward algorithm \citep{Baum1966,Rabiner1989}, or
rather by the forward part of this algorithm.  \cite{Lystig2002}
changed the forward algorithm in such a way as to allow computing the
gradients of the log-likelihood at the same time.  They start by
rewriting the likelihood as follows (for ease of exposition the
dependence on the model parameters is dropped here):
\begin{equation}
	L_{T} = Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) = \prod_{t=1}^{T} 
Pr(\vc{O}_{t}|\vc{O}_{1}, 
	\ldots, \vc{O}_{t-1}), 
	\label{condLike}
\end{equation}
where $Pr(\vc{O}_{1}|\vc{O}_{0}):=Pr(\vc{O}_{1})$. Note that for a 
simple, i.e.\ observed, Markov chain these probabilities reduce to 
$Pr(\vc{O}_{t}|\vc{O}_{1},\ldots, 
\vc{O}_{t-1})=Pr(\vc{O}_{t}|\vc{O}_{t-1})$.
The log-likelihood can now be expressed as:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log[Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, 
\vc{O}_{t-1})].
	\label{eq:condLogl}
\end{equation}

To compute the log-likelihood, \cite{Lystig2002} define the following 
(forward) recursion:
\begin{align}
	\phi_{1}(j) &:= Pr(\vc{O}_{1}, S_{1}=j) = \pi_{j} b_{j}(\vc{O}_{1}) 
	\label{eq:fwd1} \\
\begin{split}
	\phi_{t}(j) &:= Pr(\vc{O}_{t}, S_{t}=j|\vc{O}_{1}, \ldots, 
\vc{O}_{t-1}) \\
	&= \sum_{i=1}^{N} [\phi_{t-1}(i)a_{ij}b_{j}(\vc{O}_{t})] \times 
(\Phi_{t-1})^{-1},
	\label{eq:fwdt} 
\end{split} 
\end{align}
where $\Phi_{t}=\sum_{i=1}^{N} \phi_{t}(i)$. Combining 
$\Phi_{t}=Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, \vc{O}_{t-1})$, and 
equation~(\ref{eq:condLogl}) gives the following expression for the 
log-likelihood:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log \Phi_{t}.
	\label{eq:logl}
\end{equation}

\subsection{Computational considerations} From
equations~(\ref{eq:fwd1}--\ref{eq:fwdt}), it can be seen that
computing the forward variables, and hence the log-likelihood, takes
$O(Tn^{2})$ computations, for an $n$-state model and a time series of
length $T$.

\subsection{Parameter estimation}

Parameters are estimated in \pkg{depmixS4} using the EM algorithm or
through the use of a general Newton-Raphson optimizer.  The EM
algorithm however has some drawbacks.  First, it can be slow to
converge towards the end of optimization (although it is usually
faster than direct optimization at the start, so possibly a
combination of EM and direct optimization is fastest).  Second,
applying constraints to parameters can be problematic; in particular,
EM can lead to wrong parameter estimates when applying constraints.
Hence, in \pkg{depmixS4}, EM is used by default in unconstrained
models, but otherwise, direct optimization is done using \pkg{Rdonlp2}
\cite{Tamura2007,Spellucci2002}, because it handles general linear
(in)equality constraints, and optionally also non-linear constraints.

\section{Using depmixS4}

Two steps are involved in using \pkg{depmixS4} which are illustrated
below with examples:
\begin{enumerate}
	\item model specification with function \code{depmix}

	\item  model fitting with function \code{fit}
\end{enumerate}

\subsection{Example data: speed}

Throughout this manual a data set called \code{speed} is used.  It
consists of three time series with three variables: reaction time,
accuracy, and a covariate Pacc which defines the relative pay-off for
speeded and accurate responding. The participant in this experiment 
switches between fast responding at chance level and relatively 
slower responding at a high level of accuracy.

Interesting hypotheses to test are: is the switching regime symmetric?
Is there evidence for two states or does one state suffice?  Is the
guessing state actually a guessing state, i.e., is the probability
of a correct response at chance level (0.5)?

\subsection{Defining models}

A dependent mixture model is defined by the number of states, and by
the response distribution functions, and can be created with the
\code{depmix}-function as follows (see the help files for other
options):

\begin{verbatim} 
mod <- depmix(rt~1, data=speed, nstates=2)
\end{verbatim}

Above code illustrates the simplest case of a univariate time series
without covariates; besides providing the data, the only other
necessity is to specify the desired number of states. Note that the
rt's are modelled here with a Gaussian distribution as that is the
default family option in \pkg{depmixS4}. A multivariate model can be
specified by providing a list of formulae rather than a single one as
above:

\begin{verbatim} 
mod <- depmix(list(rt~1,corr~1), data=speed, nstates=2, 
family=list(gaussian(),multinomial())
\end{verbatim}

Here it is also necessary to provide the family functions for each of
the responses.  Currently, the gaussian() and multinomial() are
implemented.

The function \code{depmix} returns an object of class \code{depmix}
which has its own summary function providing the parameter values of
the model. The object consists of three main parts: the prior model,
which specifies the initial state probabilities, the transition
models, specifying the transition probabilities for each state, and
the response models, specifying the densities for each response and
each state. See the help files for further details.

Except in simple cases, starting values can be a problem in latent
Markov models, and so in general it's best to provide them if you have
a fairly good idea of what to expect.  Providing starting values is
done through three arguments: respstart, trstart, and instart, for
response related parameters, transition parameters and prior
parameters, respectively.  The use of setpars reveals the ordering of
parameters that should be used:

\begin{verbatim}
mod <- depmix(list(rt~1,corr~1), data=speed, nstates=2, 
family=list(gaussian(),multinomial())
setpars(mod,1:npar(mod))
\end{verbatim}

See the paragraph on parameter numbering below for further details.

\subsection{Transition matrix and initial state probabilities}

The transition matrix is parametrized as a list of multinomial
logistic models.  The initial state probabilities are similarly
parametrized as a multinomial logistic model.  Both models use a base
category parametrization, meaning that the parameter for the base
category is fixed at zero.  The default base category is the first
state.  Hence, for example, for a 3-state model, the initial state
probability model would have three parameters of which the first is
fixed at zero and the other two are freely estimated.

%See
%\citet{Agresti20??} for multinomial logistic models and various
%parameterizations.

Covariates can be specified using a one-sided formula as in the
following example:

\begin{verbatim}
mod <- depmix(list(rt~1,corr~1), data=speed, nstates=2,
transition=~Pacc)
\end{verbatim}

Note that this can be done for the initial state probabilities by
specifying prior=~X1, where X1 is the desired covariate.  The result
of this is that the transition probabilities are now dependent on the
covariate Pacc (which is an experimenter controlled variable to induce
switching between guessing and accurate responding).

\subsection{Fitting models}

Fitting models is done using the function \code{fit}.  The standard
call only requires a model object of class \code{depmix}:

\begin{verbatim} 
mod <- depmix(list(rt~1,corr~1), data=speed, nstates=2, 
family=list(gaussian(),multinomial())
fmod <- fit(mod)
\end{verbatim}

The function returns an object of class \code{depmix.fitted} which
extends the \code{depmix} class, adding convergence information and
possibly information about constraints if these were applied.  The
function provides some online output on the progress of the
optimization, the precise form of the output depends on the
optimization method chosen.

Class \code{depmix} and \code{depmix.fitted} have \code{logLik},
\code{AIC} and \code{BIC} methods to provide fit statistics.

\subsection{Fixing and constraining parameters}

Constraining and fixing parameters is done using the \code{conpat}
argument to the \code{depmix.fit}-function, which specifies for each
parameter in the model whether it's fixed (0) or free (1 or higher).
Equality constraints can be imposed by having two parameters have the
same number in the \code{conpat} vector.  When only fixed values are
required the \code{fixed} argument can be used instead of
\code{conpat}, with zeroes for fixed parameters and other values (ones
e.g.) for non-fixed parameters.  Fitting the models subject to these
constraints is handled by the optimization routine \code{donlp2}.

\paragraph{Parameter numbering} When using the \code{conpat} and
\code{fixed} arguments, complete parameter vectors should be supplied,
i.e., these vectors should have length of the number of parameters of
the model, which can be obtained by calling \code{npar(object)}.
Parameters are numbered in the following order:
\begin{enumerate}
	\item  the prior model parameters

	\item  the parameters for the transition models

	\item  the response model parameters per state (and subsequently
	per response in the case of multivariate time series)

\end{enumerate}

To see the ordering of parameters use the following:
\begin{verbatim}
mod <- setpars(mod, value=1:npar(mod))mod
\end{verbatim}

To see which parameters are fixed (by default only baseline parameters
in the multinomial logistic models for the transition models and the
initial state probabilities model:
\begin{verbatim}
mod <- setpars(mod,
getpars(mod,which="fixed"))mod
\end{verbatim}

\bibliography{all,ingmar}

\end{document}