### 2008-11-28 papers/individual/individual.tex
Updated individual paper with Brenda's comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@245 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%&LaTeX
\documentclass[a4paper,12pt,man]{apa} % nobf, doc, apa

%\usepackage[english]{babel} % ik krijg anders foute citaties...in Zweeds formaat???

\usepackage[]{amsmath, amsfonts, amstext, amsthm}
\usepackage{amssymb}
\usepackage[]{graphics}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{epstopdf}

\newcommand{\citep}{\cite}
\newcommand{\citet}{\citeA}
\newcommand{\mat}{\mathbf}
\newcommand{\vc}{\mathbf}

\newcommand{\pkg}{\texttt}
\newcommand{\code}{\texttt}

\renewcommand{\baselinestretch}{2}

\title{A Framework for Discrete Change}

\twoauthors{Ingmar Visser \& Brenda R. J. Jansen}{Maarten Speekenbrink}

\date{\today}

\twoaffiliations{Department of Psychology, University of Amsterdam}
{Department of Cognitive, Perceptual and Brain Sciences, University College London}

\note{Correspondence concerning this article should be addressed to:
 \\
 Ingmar Visser \\
 Department of Psychology, University of Amsterdam \\
 Roetersstraat 15 \\
 1018 WB Amsterdam \\
 phone: +31 (20) 5256723 \\
 fax: +31 (20) 6390279 \\
 email: i.visser@uva.nl}

\shorttitle{Modelling Discrete Change}

\abstract{A class of models is developed for measuring and detecting
discrete change in learning and development.  The basic model for
detecting such change is the latent or hidden Markov model.
Traditionally, these models were restricted to categorical, mostly
binary, observed variables, placing severe restrictions on possible
measurement models.  In this paper, the basic model is extended to
include various distributions for the observed variables, including
mixed multi-variate distributions.  Moreover, there is an option to
include time-varying predictors on the observed distributions.  In
effect, this model consists of mixtures of generalized linear models
with Markovian dependencies over time to model the change process.  In
addition, the transition parameters can be estimated with covariates,
such that the switching regime between states depends on
characteristics of the individual or the experimental situation.  The
model is illustrated with examples of participants' learning in the
Iowa Gambling Task and in the Weather Prediction Task.}

\begin{document}
\maketitle

% intro
Discrete change frequently occurs in learning and development: in
learning concepts and discrimination learning, in performance on
Piagetian tasks such as the conservation of liquid task and the
balance scale task, and in conditioning.  This chapter is concerned
with detecting and characterizing such discrete changes in
(individual) time series data.  We present a framework of dependent
mixture models that can be used to test whether discrete learning
events are present and, if so, to test at which point in time those
changes are taking place.  Before presenting the model in formal terms
and providing some illustrations, we first review some examples from
the developmental psychology literature in which discrete change is
typically found.

% phenomena of discrete change in development/learning
Piagetian developmental theory assumes step-wise changes in the rules
that children apply in all kinds of tasks such as in conservation
learning and in the balance scale task (SieglerAlibali19??).
\citet{Maas1992} developed a catastrophe model to describe step-wise
changes predicted to occur by Piagetian theory; importantly, they
derived precise statistical criteria for testing whether two
discretely different stages exist.  They applied the catastrophe model
to development on the conservation of liquid task (InhelderPiaget19??)
in which children have to judge relative volumes of liquid in glasses
of different heights and widths.  Young children tend to ignore the
width dimension and hence always choose the glass with the highest
level of liquid (InhelderPiaget19??).  It is assumed that there is a
sudden transition to a new rule, for example, a rule in which the
children also take the width of the glasses into account when judging
the volume of liquids.

% \citet{Maas1992} 
% opzoeken: Siegler & Alibali 19??, intro textbook
% opzoeken: Inhelder & Piaget, 19??

% balance scale rules
Jansen \& Maas (2001) applied the catastrophe model to
development of reasoning on the balance scale task (Siegler, 1981).
In the balance scale task participants have to judge which side of a
balance goes down when the number of weights and their distances to
the fulcrum are varied over trials.  Younger children tend to ignore
the distance dimension in this task, and instead focus solely on the
number of weights on each side of the fulcrum.  This strategy for
solving balance scale items is called Rule 1 \citet{Siegler1981}.  Older
children include the distance dimension in determining their response
to balance scale problems; however, they only do so when the weight
dimension does not differ between the sides of the balance, i.e., when
the number of weights is equal on both sides of the balance scale.
This strategy is called Rule 2 \cite{Siegler1981}.

% Jansen 2001: aanpassen in de paper ref ipv het proefschrift

% hysteresis on the balance scale
\citet{Jansen2001} found evidence for stage-wise transitions, or
discrete change in our model, between Rule 1 and Rule 2 by testing
criteria that were derived from the catastrophe model.  In particular,
she found bimodal test scores and inaccessibility.  The latter means
that there are no in-between strategies: children apply either Rule 1
or Rule 2 and there is no in-between option.  \citet{Jansen2001} also
found evidence for hysteresis: the phenomenon that switching between
strategies is asymmetric.  Children can switch from Rule 1 to Rule 2
and back, but this occurs at different values of a continuously
changing independent variable.  In particular, if the distance
dimension in the balance scale problems is made increasingly more
salient by increasing the distance difference between weights on
either side of the balance scale, children may switch from Rule 1 to
Rule 2.  If subsequently the distance difference is decreased again,
children may switch back to using Rule 1.  Hysteresis is the
phenomenon that this switch back occurs at a different value of the
control variable, in this case the distance difference.

% conditioning and addiction research
Also in animal learning and conditioning, evidence is found for sudden
changes in response behavior.  For example, \citet{Gallistel2004}
found evidence for sudden onset of learning: at the start of the
learning experiment, pigeons did not learn anything and performance
was stable; after a number of trials, learning kicked in and there
were large increases in performance.  \citet{Gallistel2004}
focused on modeling the distribution of onset times: that is, the
trials at which learning suddenly takes off.  A similar interest in
process onset times is found in addiction research.  For example,
\citet{Sher2004} study the age at which children start using alcohol
and how this related to eventual outcomes in terms of addiction.

% discrimination and categorization learning
Sudden transitions in learning are also observed in simple
discrimination learning paradigms in which participants learn to
discriminate a number of stimuli based on a single dimension such as
form or color.  \citet{Raijmakers2001} found evidence for two
different strategies applied by children when faced with such a
learning task.  \citet{Schmittmann2006} reanalyzed their data using
hidden Markov models to show that both strategies are characterized by
sudden transitions in the learning process.

% latent Markov versus time series analyses/outline
In above mentioned applications, the data consist mostly of a few
repeated measurements administered to large groups of participants.
The focus in the current chapter is rather on data that consist of
many repeated measurements, or time series, observed in only a few
participants or even just a single participant.  For example,
\citet{Visser2009} analyzed data from a single participant in an
experimental task that manipulates the trade-off between speed and
accuracy.  The data consisted of three time series with each around
150 repeated measurements of both reaction time and accuracy.  Below
we provide examples of analyzing time series from single participants
from two experiments; one from the Iowa Gambling Task and one from the
Weather Prediction Task.  The interest in these tasks is to show that
participants develop different strategies over time in responding to
the stimuli and that the transition from one strategy to the next is a
discrete event.  Before providing these illustrations, we give a
formalization of dependent mixture models and a brief overview of the
\pkg{depmixS4} package that was developed to specify and fit such
models.

\section{Dependent Mixture Models}

% outline
In this section we describe a class of models which are especially
suitable for describing and testing discrete change in (individual)
time series data.  The dependent mixture model is similar to, but
slightly different from, two other types of models that are in use for
modelling discrete change: the hidden and the latent Markov models.

% history of Markov models/LMM applications
Markov models have been used in the social sciences; for example, in
analyzing language learning \cite{Miller1952,Miller1963} and in the
analysis of paired associate learning \cite{Wickens1982}.  In these
models, the focus is on survey type data: a few repeated measurements
taken in a large sample.  \citeA{Langeheine1990} discuss latent Markov
models and their use in sociology and political science (see also
McCutcheon, 1987).  Latent Markov models have been used in studying
development of math skills \citep{Collins1992}, and in medical
applications \citep{Reboussin1998}; \citet{Kaplan2008} provides an
overview of such models, also called stage-sequential models in the
developmental psychology literature.

\nocite{McCutcheon1987}

% hidden Markov models
Hidden Markov models (HMMs) tend to be used in the analysis of long
univariate (individual) timeseries.  For example, HMMs are the model
of choice in speech recognition applications \cite{Rabiner1989}.  In
biology, HMMs are used to analyze DNA sequences \cite{Krogh1998} and
in econometric science, to analyze changes in stock market prices and
commodities \cite{Kim1994}.

% depmix model
The dependent mixture model (DMM) that we propose here spans the range
from latent Markov models for few repeated measurements with many
participants to hidden Markov models for individual time series.  In
addition, the dependent mixture model includes multivariate responses
and covariates.  The dependent mixture model for a time series $O_{t}$
consists of the following elements:
\begin{enumerate}
	\item $S$ is a collection of discrete states
	\item $S_{t} = \mat{A}S_{t-1}+\xi_{t}, \mat{A}$, a transition matrix
	\item $O_{t} = \mat{B}(S_{t}) + \zeta_{t}, \mat{B}$,  an observation density.
\end{enumerate}
Here $\xi_{t}$ and $\zeta_{t}$ are independent error processes 
\cite{Elliott1995}.

% state space
The state space, which is a set of discrete states, captures the
different states that the learning or developmental process under
consideration can be in.  In the balance scale example mentioned
above, children are applying one of two possible rules in responding
to the items.  The states are characterized by their corresponding
observation densities.  Using a particular rule in the balance scale
task leads to correct answers on some items and incorrect answers on
others.  A different rule may lead to correct answers on some items
and to guessing behavior on other items.  As another example, in
analyzing categorization learning data, in which participants learn to
categorize a set of objects, a typical initial state is that
participants are guessing because at the start of the task they have
no knowledge of which features are important in categorization.
Hence, the states in the state space represent knowledge states of the
participants, such that different knowledge states lead to different
observed behaviors or responses.

% transitions/matrix
Transition matrix $\mat{A}$ describes the transitions between states
over repeated measurements.  This matrix summarizes the probabilities
of transitioning from one state to another, which represents learning
or development.  The transition model contains the Markov assumption:
$$Pr(S_{t}|S_{t-1}, \ldots, S_{1}) = Pr(S_{t}|S_{t-1}),$$
where $Pr$ is the probability.  The Markov assumption means that the
current state (at time $t$) only depends on the previous state
$S_{t-1}$, and not on earlier states.

% observation densities
The observation densities $\mat{B}$ form the measurement part of the
model; these describe the distributions of the observations
conditional on the current state. Hence, these distributions
characterize the state, and in our examples, these characterize the
rule that a participant is using at a given measurement occasion.

% log-likelihood: kan dit helemaal weg???
The log-likelihood of DMMs is usually computed by the
so-called forward-backward algorithm \citep{Baum1966,Rabiner1989}, or
rather by the forward part of this algorithm.  \citet{Lystig2002}
changed the forward algorithm in such a way as to allow computing the
gradients of the log-likelihood at the same time.  They start by
rewriting the likelihood as follows (for ease of exposition the
dependence on the model parameters is dropped here):
\begin{equation}
	L_{T} = Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) = \prod_{t=1}^{T}
Pr(\vc{O}_{t}|\vc{O}_{1},
	\ldots, \vc{O}_{t-1}),
	\label{condLike}
\end{equation}
where $Pr(\vc{O}_{1}|\vc{O}_{0}):=Pr(\vc{O}_{1})$. Note that for a
simple, i.e.\ observed, Markov chain these probabilities reduce to
$Pr(\vc{O}_{t}|\vc{O}_{1},\ldots,
\vc{O}_{t-1})=Pr(\vc{O}_{t}|\vc{O}_{t-1})$.
The log-likelihood can now be expressed as:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log[Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots,
\vc{O}_{t-1})].
	\label{eq:condLogl}
\end{equation}

To compute the log-likelihood, \citet{Lystig2002} define the following
(forward) recursion:
\begin{align}
	\phi_{1}(j) &:= Pr(\vc{O}_{1}, S_{1}=j) = \pi_{j} b_{j}(\vc{O}_{1})
	\label{eq:fwd1} \\
\begin{split}
	\phi_{t}(j) &:= Pr(\vc{O}_{t}, S_{t}=j|\vc{O}_{1}, \ldots,
\vc{O}_{t-1}) \\
	&= \sum_{i=1}^{N} [\phi_{t-1}(i)a_{ij}b_{j}(\vc{O}_{t})] \times
(\Phi_{t-1})^{-1},
	\label{eq:fwdt}
\end{split}
\end{align}
where $\Phi_{t}=\sum_{i=1}^{N} \phi_{t}(i)$. Combining
$\Phi_{t}=Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, \vc{O}_{t-1})$, and
equation~(\ref{eq:condLogl}) gives the following expression for the
log-likelihood:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log \Phi_{t}.
	\label{eq:logl}
\end{equation}

Note that so far no assumptions have been made about the response
distributions $b_{j}$, hence these can be arbitrary univariate or
multivariate distributions.

\section{DepmixS4}

% outline/abstract: this has redundant information, mostly discussed
% in detail below
\pkg{depmixS4} implements a general framework for defining and fitting
dependent mixture models in the R programming language \citep{R2008}.
This includes standard Markov models, latent/hidden Markov models, and
latent class and finite mixture distribution models.  The models can
be fitted on mixed multivariate data with multinomial and/or gaussian
distributions.

% design goals
The \pkg{depmixS4} package was motivated by the fact that Markovian
models are used commonly in the social sciences, but no comprehensive
package was available for fitting such models in R. Common programs
for Markovian models include Panmark \citep{Pol1996}, and for latent
class models Latent Gold \citep{Vermunt2003}.  Those programs are
lacking a number of features that were needed in our research.  In
particular, \pkg{depmixS4} was designed to meet the following goals:
\begin{enumerate}

	\item to be able to fit transition models with covariates, i.e.,
	such that the transition probabilities depend on a time-dependent 
	covariate.

	\item to be able to include covariates in the prior or initial
	state probabilities of models.

	\item to allow for easy extensibility; in particular, to be able
	to add new response distributions, both univariate and
	multivariate, and similarly to be able to allow for the addition
	of other transition models, e.g., models for observations in 
	continuous time.

\end{enumerate}

Although \pkg{depmixS4} is designed to deal with longitudinal or time
series data, for say $T>100$, it can also handle the limit case with
$T=1$ in analyzing cross-sectional data.  In those cases, there are no
time dependencies between observed data, and the model reduces to a
finite mixture model \cite{McLachlan2000}, or a latent class model
\cite{McCutcheon1987}.  Although there are other specialized (R)
packages to deal with mixture data, these don't allow the inclusion of 
covariates on the prior probabilities of class membership.

\pkg{depmixS4} is built using S4 classes (object oriented classes in R)
to allow easy extensibility (Chambers, 1998). A depmix model consists 
of the following: 
\begin{enumerate} 
	\item A model for the initial or prior probabilities 
	\item A list of transition models; one model for each row of the 
	transition matrix
	\item A list of response models for each of the states of the model; 
	this is a list of lists in the case of multivariate responses. 
\end{enumerate} 
Each of these are adressed briefly below.

\subsection{Transition and prior probabilities models}

% transition and initial state probs
By default, each row of the transition matrix and the vector of
initial state probabilities is modelled as a baseline-category
multinomial logistic model \cite<see>[chapter 7]{Agresti2002}.  This
means that covariates can be used as predictors in these models.  In
particular, the model for each multinomial is:
\begin{equation} 
	\pi_{i} = \frac{\exp(\alpha_{i}+\mathbf{\beta}^{`}_{i}\vc{x})}
				{1+\sum_{j=1}^{J-1}\exp(\alpha_{j}+\mathbf{\beta}^{`}_{j}\vc{x})},
\end{equation}
where $\pi_{i}$ is the probability (e.g.\ the probability of the
$i$-th initial state); the $\alpha_{i}$ are the category intercepts;
the $\mathbf{\beta}_{i}$ are the category regression coefficients;
$\vc{x}$ is a vector of covariates or predictors; $J$ is the number of
states; in this example, $J$ serves as the baseline-category, meaning
that $\alpha_{J}$ and $\mathbf{\beta}_{J}$ are zero.

\subsection{Response models}

% response models
Response models in \pkg{depmixS4} interface with the \code{glm} 
functions available in R and with the \pkg{nnet}-package for the 
multinomial logistic models. Normally distributed data can hence be 
modelled with direct effects included as well. For example, normal 
data are modelled as:
\begin{equation}
	O_{t}|S_{t} = \mu + \mathbf{\beta}^{`}\vc{x} + \epsilon_{t},
\end{equation}
where $\mu$ is the (state-dependent) mean; $\vc{x}$ is a vector of
covariates or predictors; $\mathbf{\beta}$ is the vector of
(state-dependent) regression coefficients; $\epsilon_{t}$ is a
normally distributed error term with (state-dependent) standard
deviation $\sigma$.

Multinomial responses are modeled by the same multinomial 
baseline-category logistic model as is used for the transition 
probabilities and the initial state probabilities. Other response 
functions from the generalized linear model family that can be used 
include: binomial logistic regression, poisson regression, gamma and 
inverse gaussian. There is also support for the multivariate normal 
distribution.

% MAARTEN: de multinomial logistic models maken weer gebruik van andere
% packages toch, kun je daar de referentie van geven?

% referentie: nnet

\subsection{Parameter Estimation}

Parameters are estimated in \pkg{depmixS4} using the EM algorithm or
through the use of a general Newton-Raphson optimizer.  EM is used by
default in unconstrained models, but otherwise direct optimization is
done using \pkg{Rdonlp2} \cite{Tamura2007,Spellucci2002}, because it
handles general linear (in)equality constraints, and optionally also
non-linear constraints.

\section{Illustrations}

Two illustrations are provided below of models that analyze single
participant time series data from two common experimental paradigms.
In both of these, participants learn different strategies through
trial and error.

\subsection{Iowa gambling task}

% what is the IGT?
The Iowa gambling task (IGT) is an experimental paradigm designed to
mimic real-life decision-making situations \cite{Bechara1994}, in the
way that it factors uncertainty, reward and punishment
\cite{Dunn2006}.  The task requires the selection of cards from four
decks.  Each deck is characterized by a certain amount of gain
(delivered on each draw), frequency of loss, and amount of loss.  Two
decks (A and B) yield consistently high rewards, but also high,
probabilistic penalties and are both (equally) disadvantageous in the
long run.  The other two decks (C and D) yield consistently smaller
rewards, but also low, probabilistic penalties and are both (equally)
advantageous in the long run.  It is assumed that the ventromedial
prefrontal cortex (VMPFC) is active in the IGT as VMPC patients show
impaired task performance.  Their preference for the decks with
immediate high rewards indicates ``myopia for the future''
\cite{Bechara1994}.

% what is the HDT? developmental trends and relevance?
\citet{Crone2004} designed a developmentally appropriate analogue of
the IGT, the Hungry Donkey Task (HDT), with a similar win and loss
schedule although the absolute amounts were reduced by a factor of 25.
The HDT is a pro-social game inviting the player to assist a hungry
donkey to collect as many apples as possible, by opening one of four
doors.  Again, doors A and B are characterized by a high constant gain
(10 apples), whereas doors C and D deliver a low constant gain (2
apples).  At doors A and C, a loss of 10 apples (A) or 2 apples (C)
is delivered in 50\% of the trials.  For doors B and D, frequency of
loss is only 10\%.  The median loss of doors B and D is 50 and 10,
respectively.  \citet{Crone2004} administered the HDT to children from
four age groups (6-9, 10-12, 13-15, and 18-25 year-olds) and concluded
that children have difficulty to take future consequences into account.

% strategic reanalysis of the HDT: different strategies found by
% finite mixture analysis
A reanalysis of this dataset \cite{Huizenga2007b} indicated that
participants might solve the task by sequentially considering the
three dimensions (constant gain, frequency of loss, and amount of
loss) in order to choose a door.  Most of the youngest children seem
to focus frequency of loss, resulting in equal preference for doors B
and D. It's assumed that frequency of loss is the dominant dimension
in the task.  Older participants seem to use a two-dimensional rule
where participants first focus on the frequency of loss and then
consider amount of loss, resulting in a preference for door D. A third
very small subgroup seems to use an integrative rule where
participants combine all three dimensions in the appropriate way.
Participants using the integrative rule pick cards from doors C and D,
which are advantageous in the long run.

% Check ref Huizenga 2007b

% problematic aspects of standard analysis
\citet{Huizenga2007b} used the last 60 trials in a series of 200
trials, assuming that behavior has stabilized after 140 trials of
learning; this could very well be wrong and it is highly likely that
there are individual differences in this learning process.  Hence,
here we analyze IGT learning data from a single participant with the
aim to establish 1) at which trial learning stops, i.e., at which
trial behavior has stabilized, and 2) what strategies participants use
during and at the end of learning.  Models with an increasing number
(from 2 through 5) of latent states were fitted to the time series of
various participants.  The responses were fitted with multinomial
logistic models for the four possible choices that participants make in
this task.  There were no covariates on any of the parameters.  The
only constraint imposed on the models' parameters was that there were
designated begin and end states.  This means that the initial state
probabilities were fixed to one for the first state and to zero for
the remaining states.  For the transition matrix this means that the
final state was an absorbing state: the probability of transitioning
out of that state is zero.  This was done to ensure that there would
be a final state in which participants end which provides the
possibility of immediately seeing participants' behavior at the end of
the task.  Models were selected using the Akaike Information Criterion
(AIC; Akaike, 1973).

\nocite{Akaike1973} % check the year!!

The data were best described by a 4-state model.  This model's
transition matrix is:
$$
\mat{A} = \begin{pmatrix} 
				0.64 & 0.33 & 0.00 & 0.03 \\
				0.10 & 0.89 & 0.01 & 0.00 \\
				0.05 & 0.00 & 0.95 & 0.00 \\
				0.00 & 0.00 & 0.00 & 1.00
		  \end{pmatrix}
$$
As can be seen from the transition matrix the final state is
absorbing.  States 2 and 3 are also fairly stable with high diagonal
probabilities of 0.89 and 0.95 respectively.  The response parameters
are presented in Table~\ref{tab:igt4} below.

% Table refs checken

\begin{table}
	\caption{Estimates for the Iowa Gambling Task model}
	\label{tab:igt4}
	\begin{tabular}{lcccccccc} \hline
		state & p(A) & p(B) & p(C) & p(D) \\
		$S_1$ & 0    & 0.06 & 0.31 & 0.63 \\
		$S_2$ & 0.33 & 0.67 & 0    & 0 \\ 
		$S_3$ & 0    & 0    & 1    & 0 \\
		$S_4$ & 0    & 0.06 & 0.94 & 0 
	\end{tabular}
\end{table}

The states have clear interpretations.  States 3 and 4 are both
dominated by C responses, and state 4 has a low probability of B
responses.  States 1 and 2 are dominated by C and D, and by A and B
responses, respectively.  To get a clearer interpretation of these
states it is necessary to consider when they are visited during the
learning process.  The Viterbi algorithm \citep{Rabiner1989} provides
us with the posterior state sequence, i.e., the sequence of states
that the participant is in at each trial.  This sequence is depicted
in Figure~\ref{fig:post4}.

% figure refs checken

\begin{figure}
\begin{center}
	\includegraphics[width=\textwidth]{graphs/post4.pdf}
	\label{fig:post4}
	\caption{Posterior state sequence for the 4 state model.}
\end{center}
\end{figure}

At the start of training there are many switches between states 1 and 
2, indicating that behavior is mostly random choices between all four 
categories, albeit with a slight preference for B and D over A and C
choices. This preference indicates a focus on the frequency of loss
associated with each of the doors as B and D both have low frequency of
loss. After this, there is a period of stable state 2 behavior
associated with responses A and B. After that there is stable state 3 
behavior, consisting only of C responses, a short transitional period 
and then state 4 behavior consisting of 94 \% C choices and some B
choices, i.e., almost optimal behavior. The choice proportions and the
model predicted choice proportions are depicted in
Figure~\ref{fig:igt4}.

\begin{figure}
\begin{center}
	\includegraphics[width=\textwidth]{graphs/igt4.pdf}
	\label{fig:igt4}
	\caption{Data and model predicted proportions of door choices in
	the IGT.}
\end{center}
\end{figure}

The initial preference for B and D choices confirms the theory
expressed in \citet{Huizenga2007} that frequency of loss is the
dominant dimension in the IGT. The final states with mostly C 
choices represents one of the optimal strategies, which consists of
both C and D responses. Note that both C and D responses generate equal
profits in the long run.

\subsection{Weather prediction task}

%Still to add models and data.

The Weather Prediction Task \cite<WPT>{Knowlton1994} %(WPT, Knowlton, Squire \& Gluck, 1994) 
is a probabilistic categorization task in which participants learn to
predict (or categorize) the state of the weather (sun or rain) on the
basis of four ``tarot'' cards (cards with abstract geometrical
patterns).  On a given trial, one, two, or three cues are present.
There are a total of 14 possible cue patterns, and each cue pattern is
associated with a particular probability distribution over the states
of the weather.  In order to perform well in the task, participants
must predict the weather in accordance with these conditional
probabilities.

%The WPT has been popular in neuropsychological research,
%particularly because amnesic patients perform this task rather
%well, despite not being able to remember actually many aspects
%of the task (or in some cases, even performing the task at all).
%This has led to the conclusion that probabilistic category learning
%depends on implicit memory, which is separate from explicit
%memory. While this conclusion is debatable (Speekenbrink, Channon
%\& Shanks, 2008), the finding of relatively unimpaired performance
%by amnesic individuals remains striking.

There are different accounts of probabilistic category learning
\citep<see e.g.>[for an overview]{Ashby2005}.  According to instance or
exemplar learning theories, participants learn by storing each
encountered cue-outcome pair.  When presented with a probe cue
pattern, a response is made by retrieving these exemplars are
retrieved from memory and weighting them according to their similarity
to the probe cue pattern.  According to associative theories,
participants learn by gradually associating the individual cues (or
cue patterns in configural learning) to the outcomes.  Finally, in
rule-learning, participants are taken to extract rules by which to
categorize the different cue patterns.  \citet{Gluck2002} proposed a
number of such rules (or strategies).  The rules differ in the way the
cues are combined into a response.  A main difference is whether
responses are based on the presence/absence of a single cue, or
whether information from multiple cues is integrated.
\citet{Gluck2002} formulated all strategies in a deterministic
and optimal manner (e.g., the multi-cue strategy corresponded to
giving the optimal response to each cue pattern).  \citet{Meeter2006}
allowed for probabilistic responding (a small probability of
responding non-optimally) as well as switches between strategies
during learning.  However, both strategies and response probabilities
remained predefined.  Alternative non-strategy based analyses of the
WPT \cite{Lagnado2006,Speekenbrink2008} have estimated response
distributions through logistic regression, allowing the regression
coefficients to change smoothly over time.

%associative, rule-based.
Here, we combine the regression and strategy-based approaches, and
analyse the behavior of a single individual performing the WPT for 200
trials.  We are particularly interested in evidence for strategy
switching and whether a DMM can recover strategies in accordance with
\citet{Gluck2002}.  We chose to analyse the ``average'' participant
(the participant with performance closest to the group average) in a
large unpublished dataset.  We let each state be characterized by a
GLM with a Binomial distributed response and logistic link function
(i.e., a logistic regression model).  The regression coefficients of a
model relating the cues to the state of the weather are given in
column ```validity'' of Table~\ref{tab:WPT}.  Note that indentical
coefficients for the model relating the cues to responses would
indicate probability matching.  A maximizing strategy is indicated by
more extreme regression coefficients.

As we fitted a DMM to the data of a single subject, it was necessary
to place some constraints on the model.  Specifically, we constrained
the state transitions to be in a ``left-right'' format (states can
only proceed to the immediately adjacent state and never back, and
must start in the initial state).  We fitted a single, two and three
state model to the data.  This showed that a two state model was
better than a single or three state model.

\begin{table}
\caption{Estimates for the weather prediction task}
\label{tab:WPT}
\begin{tabular}{lcccccccccc} \hline
 & & \multicolumn{1}{c}{} && \multicolumn{1}{c}{1 state} & & \multicolumn{2}{c}{2 state} &&
 \multicolumn{2}{c}{2 state (constr.)} \\ \cline{5-5} \cline{7-8} \cline{10-11}
parameter && validity && $S_1$ & & $S_1$ & $S_2$ & & $S_1$ & $S_2$ \\ \hline
(intercept) && 0 && -0.69 & & -2.73 & 0.88 & & -1.24 & 0 \\
cue 1 && 2.10 && 1.69 && 2.12 & 1.60 && 1.65 & 1.97 \\
cue 2 && 0.58 && 1.12 && 0.97 & 1.63 && 0 & 1.92 \\
cue 3 && -0.58 && -0.49 && 0.91 & -2.03 && 0 & -1.58 \\
cue 4 && -2.10 && -1.32 && 0.69 & -3.16 && 0 & -2.67 \\ \hline
 & & & & \multicolumn{1}{c}{AIC=204.47} & & \multicolumn{2}{c}{AIC=187.50} &&
 \multicolumn{2}{c}{AIC=185.24}
\end{tabular}
\end{table}

Investigation of the parameter estimates (see Table~\ref{tab:WPT})
showed that the regresion coefficient for the first cue was of much
larger magnitude than that of the other cues. As such, the first
state seems representative of single cue strategy.  Alternatively, as
all regression coefficients are positive, it could indicate a
``counting'' heuristic, where the propensity of ``sun'' responses
increases when more cues are present (regardless of which cues they
are).  However, in that case, we would expect the regression
coefficients to be of roughly identical magnitude. The second state
seemed to represent a multi-cue strategy, as all cues had regression
coefficients of reasonable magnitude in the direction of the objective
cue validities.

To reduce the degrees of freedom and improve parameter estimates, we
implemented constraints to force state 1 into a single cue strategy
(fixing the coefficients of the remaining three cues to 0) and state 2
in a multi-cue strategy (forcing the intercept to 0).  These
restrictions resulted in a better AIC value of AIC=185.24 (df=7).
Interestingly, the single cue strategy was somewhat different than
described by \citet{Gluck2002}.  Parameter estimates indicated relatively
more consistent predictions of ``rain'' in the absence of cue 1
($Pr(\text{sun}) = 0.22$) and more inconsistent predictions of ``sun''
in the presence of cue 1 ($Pr(\text{sun}) = 0.60$).  The cue weights
of the multi-cue strategy were in the direction of the optimal
weights.  The Viterbi state sequence indicated that the participant
used the single cue strategy for the first 60 trials, and then
switched to the multi-cue strategy.

\section{Discussion}

General issues
\begin{itemize}
	\item It is feasible to fit hidden Markov models in moderate length individual time series
	\item illustrations show results that are consistent with usual group based analyses AND 
	provide extra information on switching between strategies over the course of learning
	\item Many applications in experimental psychology: examples????
\end{itemize}

Specific insights from the illustrations: suggestions welcome

% IGT
confirmation of existence of different strategies

dynamics of learning now known

different strategies/rules observed during learning (eg AB and 
guessing strategies)

% WPT
Analyzing the WPT data on an individual level allowed us to precisely
estimate idiographic strategies and their progression during learning.
Moreover, using DMMs, we can combine an individual and group level
analysis, increasing reliability of the individual analyses whilst
allowing for substantial individual differences.  Previous analyses
\cite<e.g.,>{Meeter2006} pre-defined individual strategies, based on
participants' self-reports in \citet{Gluck2002}.  Interestingly,
\citet{Gluck2002} noted that these self-reports often did not
correspond to the strategies evident in participants' responses.  It
is well known that people may find it difficult to verbalize the way
in which they integrate multiple sources of information.  As such,
estimating strategies directly from participants' responses will
result in more valid assessment of their response strategies, and
provides a direct test of the validity of the strategy set identified
by \citet{Gluck2002}.

Possible future extensions in depmixS4
\begin{enumerate}
	\item richer measurement models, eg factor models, AR models etc
	\item richer transition models, eg continuous time measurement occasions
	\item explicit state durations
	\item identifiability of models
	\item model selection
\end{enumerate}

\section*{Author note}

Ingmar Visser was supported by an EC Framework 6 grant, project 516542
(NEST).  Maarten Speekenbrink was supported by the ESRC Centre for
Economic Learning and Social Evolution (ELSE).  Thanks to Hilde
Huizenga, Brenda Jansen and Anna van Duijvenvoorde for the Iowa
Gambling task data and David Lagnado for the Weather Prediction Task
data.

\bibliography{all,ingmar,individual}

\end{document}