### 2009-07-08 papers/jss/depmix-manual.Rnw
Added .Rnw file of the jss paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@280 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

% -*- mode: noweb; noweb-default-code-mode: R-mode; -*-
%&LaTeX

\documentclass[]{report}

\usepackage[]{amsmath, amsfonts, amstext, amsthm}

\batchmode

\usepackage{Sweave}

\usepackage{Rd}

\usepackage[round]{natbib}

\title{depmix: An R-package for fitting mixture models on mixed
multivariate data with Markov dependencies}

\author{Ingmar Visser\thanks{
Correspondence concerning this manual should be adressed to:
Ingmar Visser, Department of Psychology, University of Amsterdam,
Roetersstraat 15, 1018 WB, Amsterdam, The Netherlands}\\
Department of Psychology, University of Amsterdam\\
i.visser@uva.nl}

\date{\today}

%%\VignetteIndexEntry{depmix: An R-package for fitting mixture models on mixed multivariate data with Markov dependencies}
%%\VignetteDepends{depmix}
%%\VignetteKeywords{R, latent/hidden Markov model, latent class model, finite mixture model, direct optimization, gradients}
%%\VignettePackage{depmix}

\SweaveOpts{prefix.string=graphs/depmix}

\newcommand{\vc}{\mathbf}
\newcommand{\mat}{\mathbf}

% \newcommand{\pkg}{\texttt}
% \newcommand{\code}{\texttt}

\begin{document}

\bibliographystyle{plainnat}

\maketitle

\begin{abstract}

	\pkg{depmix} implements a general class of mixture models with
	Markovian dependencies between them in the R programming language
	\citep{R2007}.  This includes standard Markov models, latent/hidden
	Markov models, and latent class and finite mixture distribution models.
	The models can be fitted on mixed multivariate data with multinomial
	and/or gaussian distributions.  Parameters can be estimated subject to
	general linear constraints, and with optional inclusion of regression
	on (time-dependent) covariates.  Parameters estimation is done through
	a direct optimization approach with gradients using the \code{Rdonlp2}
	optimization routine.  A number of illustrative examples are included.

\end{abstract}

\tableofcontents

\chapter{Introduction}

Markov and latent Markov models are frequently used in the social
sciences, in different areas and applications.  In psychology, they
are used for modelling learning processes, see \citet{Wickens1982}, for an
overview, and \citet{Schmittmann2006} for a recent application.  In
economics, latent Markov models are commonly used as regime switching
models, see e.g.\ \citet{Kim1994} and \citet{Ghysels1994}.  Further
applications include speech recognition \citep{Rabiner1989}, EEG analysis
\citep{Rainer2000}, and genetics \citep{Krogh1998}.  In those latter areas
of application, latent Markov models are usually referred to as hidden
Markov models.

The \pkg{depmix} package was motivated by the fact that Markov models
are used commonly in the social sciences, but no comprehensive package
was available for fitting such models.  Common programs for Markovian
models include Panmark \citep{Panmark3-96}, and for latent class models
Latent Gold 
\citep{LatentGold3-2003}.  Those programs are lacking a number of
important features, besides not being freely available.  In
particular, \pkg{depmix}: 1) handles multiple case, or multiple group,
analysis; 2) handles arbitrarily long time series; 3) estimates models
with general linear constraints between parameters; 4) analyzes mixed
distributions, i.e., combinations of categorical and continuous
observed variables; 5) fits mixtures of latent Markov models to deal
with population heterogeneity; 6) can fit models with covariates.
Although \pkg{depmix} is specifically meant for dealing with
longitudinal or time series data, for say $T>100$, it can also handle
the limit case with $T=1$.  In those cases, there are no time
dependencies between observed data, and the model reduces to a finite
mixture model, or a latent class model.  In the next chapter, an outline
is provided of the model and the likelihood equations.  In the
chapters after that a number of examples are presented.

\section*{Acknowledgements}

I am indebted to many people for providing help in writing this package.
First and foremost Maartje Raijmakers and Verena Schmittmann tested
countless earlier versions, spotted bugs and suggested many features.
Moreover, Maartje Raijmakers provided the discrimination data set.  Han van
der Maas provided the speed-accuracy data and thereby neccessitated
implementing models with time-dependent covariates.  Conor Dolan and Raoul
Grasman both provided valuable advice on statistics in general and
optimization in particular.

\chapter{Dependent mixture models}

The data considered here, has the general form $O_{1}^{1}, \ldots,
O_{1}^{m}$, $O_{2}^{1}, \ldots, O_{2}^{m}$, \ldots, $O_{T}^{1}, \ldots,
O_{T}^{m}$ for an $m$-variate time series of length $T$.  As an
example, consider a time series of responses generated by a single
subject in a reaction time experiment.  The data consists of three
variables, reaction time, accuracy and a covariate which is a pay-off
factor which determines the reward for speed and accuracy.  These
variables are measured at 168, 134 and 137 occasions respectively.
Below, a summary is provided for these data, as well as a plot of the
first timeseries, which is selected by \code{nind=1}.

<<echo=FALSE, results=hide>>=
.libPaths(new="/Users/ivisser/Library/R/library/")
library(depmix)
load("models.Rda")
@

<<>>=
data(speed)
summary(speed)
@
\begin{figure}[htbp]
  \begin{center}
<<fig=TRUE>>=
plot(speed,nind=1) 
@
	\caption{Reaction times, accuracy and pay-off values for
	the first series of responses in dataset \texttt{speed}.}
  \end{center}
\end{figure}

The latent Markov model is commonly associated with data of this type,
albeit usually only multinomial variables are considered.  However,
common estimation procedures, such as those implemented in
\citet{Panmark3-96} are not suitable for long time series due to
underflow problems.  In contrast, the hidden Markov model is typically
only used for `long' univariate time series.  In the next section, the
likelihood and estimation procedure for the hidden Markov model is
described, given data of the above form.

The dependent mixture model is defined by the following elements:
\begin{enumerate}

	\item a set $\vc{S}$ of latent classes or states $S_{i},\, i=1,
	\ldots , n$,

	\item a matrix $\mat{A}$ of transition probabilities $a_{ij}$ for
	the transition from state $S_{i}$ to state $S_{j}$,

	\item a set $\vc{B}$ of observation functions $b_{j}(\cdot)$ that
	provide the conditional probabilities associated with latent state
	$S_{j}$,

	\item a vector $\pmb{\pi}$ of latent state initial probabilities
	$\pi_{i}$
\end{enumerate}
When transitions are added to the latent class model, it is more 
appropriate to refer to the classes as states. The word class is 
rather more associated with a stable trait-like attribute whereas a 
state can change over time.

\section{Likelihood}

The loglikelihood of hidden Markov models is usually computed by the
so-called forward-backward algorithm \citep{Bau66,Rab89}, or rather by
the forward part of this algorithm.  \cite{Lys2002} changed the forward
algorithm in such a way as to allow computing the gradients of the
loglikelihood at the same time.  They start by rewriting the
likelihood as follows (for ease of exposition the dependence on the
model parameters is dropped here):
\begin{equation}
	L_{T} = Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) = \prod_{t=1}^{T} 
Pr(\vc{O}_{t}|\vc{O}_{1}, 
	\ldots, \vc{O}_{t-1}), 
	\label{condLike}
\end{equation}
where $Pr(\vc{O}_{1}|\vc{O}_{0}):=Pr(\vc{O}_{1})$. Note that for a 
simple, i.e.\ observed, Markov chain these probabilities reduce to 
$Pr(\vc{O}_{t}|\vc{O}_{1},\ldots, 
\vc{O}_{t-1})=Pr(\vc{O}_{t}|\vc{O}_{t-1})$.
The log-likelihood can now be expressed as:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log[Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, 
\vc{O}_{t-1})].
	\label{eq:condLogl}
\end{equation}

To compute the log-likelihood, \cite{Lys2002} define the following 
(forward) recursion:
\begin{align}
	\phi_{1}(j) &:= Pr(\vc{O}_{1}, S_{1}=j) = \pi_{j} b_{j}(\vc{O}_{1}) 
	\label{eq:fwd1} \\
\begin{split}
	\phi_{t}(j) &:= Pr(\vc{O}_{t}, S_{t}=j|\vc{O}_{1}, \ldots, 
\vc{O}_{t-1}) \\
	&= \sum_{i=1}^{N} [\phi_{t-1}(i)a_{ij}b_{j}(\vc{O}_{t})] \times 
(\Phi_{t-1})^{-1},
	\label{eq:fwdt} 
\end{split} 
\end{align}
where $\Phi_{t}=\sum_{i=1}^{N} \phi_{t}(i)$. Combining 
$\Phi_{t}=Pr(\vc{O}_{t}|\vc{O}_{1}, \ldots, \vc{O}_{t-1})$, and 
equation~(\ref{eq:condLogl}) gives the following expression for the 
log-likelihood:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log \Phi_{t}.
	\label{eq:logl}
\end{equation}

The above forward recursion can readily be generalized to mixture 
models, in which it is assumed that the data are realizations of a 
number of different LMMs and the goal is to assign posterior 
probabilities to sequences of observations. This situation occurs, 
for example, in learning data where different learning strategies may 
lead to different answer patterns. From an observed sequence of 
responses, it may not be immediately clear from which learning 
process they stem. Hence, it is interesting to consider a mixture of 
latent Markov models which incorporate restrictions that are 
consistent with each of the learning strategies.

To compute the likelihood of a mixture of $K$ models, define the 
forward recursion variables as follows (these variables now have an 
extra index $k$ indicating that observation and transition 
probabilities are from latent model $k$):
\begin{align}
\begin{split}
\phi_{1}(j_{k}) &=  Pr(\vc{O}_{1}, 
S_{1}=j_{k})=p_{k}\pi_{j_{k}}b_{j_{k}}(\vc{O}_{1}).
\end{split}\label{eq:fwd1mix} \\
\begin{split}
\phi_{t}(j_{k})   &=   Pr(\vc{O}_{t}, S_{t}=j_{k}|\vc{O}_{1}, \ldots, 
\vc{O}_{t-1}) \\
			&= \left[ \sum_{k=1}^{K} \sum_{i=1}^{n_{k}} \phi_{t-1}(i_{k}) 
			a_{ij_{k}}b_{j_{k}}(\vc{O}_{t}) \right] \times (\Phi_{t-1})^{-1},
\end{split}\label{eq:fwdtmix} 
\end{align}
where $\Phi_{t} = \sum_{k=1}^{K}\sum_{i=1}^{n_{k}} \phi_{t}(j_{k})$.
Note that the double sum over $k$ and $n_{k}$ is simply an enumeration
of all the states of the model.  Now, because $a_{ij_{k}}=0$ whenever
$S_{i}$ is not part of component $k$, the sum over $k$ can be dropped
and hence equation~\ref{eq:fwdtmix} reduces to:
\begin{equation}
	\phi_{t}(j_{k}) = \left[ \sum_{i=1}^{n_{k}} \phi_{t-1}(i_{k}) 
			a_{ij_{k}}b_{j_{k}}(\vc{O}_{t}) \right] \times (\Phi_{t-1})^{-1}
\end{equation}
The loglikelihood is computed by applying equation~\ref{eq:logl} on
these terms.  For multiple cases, the log-likelihood is simply the sum
over the individual log-likelihoods.

\paragraph{Computational considerations} From
equations~(\ref{eq:fwd1}--\ref{eq:fwdt}), it can be seen that
computing the forward variables, and hence the log-likelihood, takes
$O(Tn^{2})$ computations, for an $n$-state model and a time series of
length $T$.  Consider a mixture of two components, one with two states
and the other with three states.  Using
equations~(\ref{eq:fwd1}--\ref{eq:fwdt}) to compute the log-likelihood
of this model one needs $O(Tn^{2})=O(T\times 25)$ computations whereas
with the mixture equations~(\ref{eq:fwd1mix}--\ref{eq:fwdtmix}),
$\sum_{n_{i}} O(n_{i}^{2}T)$ computations are needed, in this case
$O(T \times 13)$.  So, it can be seen that in this easy example the
computational cost is almost halved.

\section{Gradients}

\newcommand{\fpp}{\frac{\partial} {\partial \lambda_{1}}}

See equations 10--12  in \cite{Lys2002} for the score recursion 
functions of the hidden Markov model for a univariate time series. 
Here the corresponding score recursion for the multivariate mixture 
case are provided. The $t=1$ components of this score recursion are 
defined as (for an arbitrary parameter $\lambda_{1}$):
\begin{align}
\psi_{1}(j_{k};\lambda_{1}) &:=  \fpp Pr(\vc{O}_{1}|S_{1}=j_{k}) \\
\begin{split} 
	&= \left[  \fpp p_{k} \right] \pi_{j_{k}}b_{j_{k}}(\vc{O}_{1}) + 
	p_{k}\left[ \fpp \pi_{j_{k}} \right] b_{j_{k}}(\vc{O}_{1}) \\
	& \qquad  + p_{k}\pi_{j_{k}} \left[ \fpp 
b_{j_{k}}(\vc{O}_{1})\right],
\end{split} \label{eq:psi1}
\end{align}
and for $t>1$ the definition is:
\begin{align}
\psi_{t}(j_{k};\lambda_{1})  & =  \frac{\fpp Pr(\vc{O}_{1}, \ldots, 
\vc{O}_{t}, S_{t}=j_{k})}
			{Pr(\vc{O}_{1}, \ldots, \vc{O}_{t-1})}  \\
\begin{split} 
	& =  
			 \sum_{i=1}^{n_{k}} \Bigg\{ \psi_{t-1}(i;\lambda_{1})a_{ij_{k}} 
			 b_{j_{k}}(\vc{O}_{t}) \\ 
			 &\qquad +\phi_{t-1}(i) \left[ \fpp a_{ij_{k}} \right] b_{j_{k}} 
(\vc{O}_{t}) \\
			&\qquad +\phi_{t-1}(i)a_{ij_{k}}  \left[ \fpp b_{j_{k}} 
(\vc{O}_{t}) \right] \Bigg\} 
			\times (\Phi_{t-1})^{-1}.
\end{split} \label{eq:psit}
\end{align}

Using above equations, \cite{Lys2002} derive the following equation 
for the partial derivative of the likelihood:
\begin{equation}
	\fpp l_{T}= 	
		\frac{\mathbf{\Psi}_{T}(\lambda_{1})}{\mathbf{\Phi}_{T}},
\end{equation}
where $\Psi_{t}=\sum_{k=1}^{K} \sum_{i=1}^{n_{k}} 
\psi_{t}(j_{k};\lambda_{1})$. 
Starting from the equation from the logarithm of the likelihood, this 
is easily seen to be correct: 
\begin{eqnarray*}
	\fpp \log Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) &=& Pr(\vc{O}_{1}, 
\ldots, \vc{O}_{T})^{-1} 
	\fpp Pr(\vc{O}_{1}, \ldots, \vc{O}_{T}) \\
	&=&  \frac{Pr(\vc{O}_{1}, \ldots, \vc{O}_{T-1})}{Pr(\vc{O}_{1}, 
\ldots, \vc{O}_{T})}  \Psi_{T} (\lambda_{1}) \\
	&=&  \frac{\mathbf{\Psi}_{T}(\lambda_{1})}{\mathbf{\Phi}_{T}}.
\end{eqnarray*}

Further, to actually compute the gradients, the partial derivatives of
the parameters and observation distribution functions are neccessary,
i.e., $\fpp p_{k}$, $\fpp \pi_{i}$, $\fpp a_{ij}$, and $\fpp
\vc{b}_{i}(\vc{O}_{t})$.  Only the latter case requires some
attention.  We need the following derivatives $\fpp
\vc{b}_{j}(\vc{O}_{t})=\fpp \vc{b}_{j}(O_{t}^{1}, \ldots, O_{t}^{m})$, for
arbitrary parameters $\lambda_{1}$. To stress that $\vc{b}_{j}$ is a
vector of functions, we here used boldface. First note that because of local
independence we can write:
\begin{equation*}
	\fpp \left[ b_{j}(O_{t}^{1}, \ldots, O_{t}^{m}) \right] = \frac{\partial} 
{\partial \lambda_{1} } \left[ b_{j}(O_{t}^{1}) \right] \times  
\left[ b_{j}(O_{t}^{2}) \right], \ldots,  \left[ b_{j}(O_{t}^{m}) \right].  
\end{equation*}
Applying the chain rule for products we get:
\begin{equation}
	\fpp [b_{j}(O_{t}^{1}, \ldots, O_{t}^{m})] =
	\sum_{l=1}^{m} \left[ \prod_{i=1, \ldots, \hat{l}, \ldots, m} 
b_{j}(O_{t}^{i}) \right] \times
	\fpp  [b_{j}(O_{t}^{l})],
	\label{partialProd}
\end{equation}
where $\hat{l}$ means that that term is left out of the product. 
These latter terms, $\frac{\partial} {\partial \lambda_{1} }  
[b_{j}(O_{t}^{k})]$, are easy to compute given either multinomial or 
gaussian observation densities $b_{j}(\dot)$

\section{Parameter estimation}

Parameters are estimated in \pkg{depmix} using a direct optimization
approach instead of the EM algorithm which is frequently used for this type
of model.  The EM algorithm however has some drawbacks.  First, it can be
slow to converge.  Second, applying constraints to parameters can be
problmatic.  The EM algorithm can sometimes lead to incorrect estimates
when constraints are applied to parameters in the M-step of the algorithm.
The package was designed to be used with the \code{npsol}-library, the
main reason being that it handles general linear (in-)equality
constraints very well. Unfortunately, \code{npsol} is not freeware and
hence is not distributed with \pkg{depmix}. Two other options are
available for optimization using  \code{nlm} and \code{optim}
respectively. Linear equality constraints are fitted through
reparametrization. Inequality constraints are fitted through adding a 
penalty to the likelihood depending on the amount by which a
constraint is not satisfied. The argument \code{vfactor} to the
fitting function can be used to control this bahavior. See details of 
this in the chapter on fitting models.

\chapter{Using depmix}

Three steps are involved in using \pkg{depmix} which are illustrated
below with examples:
\begin{enumerate}
	\item  data specification with function \code{markovdata}

	\item  model specification with function \code{dmm}

	\item  model fitting with function \code{fitdmm}
\end{enumerate}

To be able to fit models, data need to in a specific format created
for this package.  Basically, data should be in the form of a matrix
with each row corresponding to measures taken at a single measurement
occasion for a single subject.  The function \code{markovdata} further
only requires one argument providing the itemtypes, being one of
categorical, continuous or covariate.  A markovdata object is a matrix
with a number of attributes.

\section{Creating data sets}

As an example we make a dataset with two variables measured at two
times 50 occasions.

<<>>=
x=rnorm(100,10,2)
y=ifelse(runif(100)<0.5,0,1)
z=matrix(c(x,y),100,2)
md=markovdata(z,itemtypes=c("cont","cat"),ntimes=c(50,50))
md[1:10,]
@

In the example below, we split the dataset \code{speed} into three
separate datasets, which we later use as an example to do multi-group
analysis.

<<>>=
data(speed)
r1=markovdata(dat=speed[1:168,],itemt=itemtypes(speed))
r2=markovdata(dat=speed[169:302,],itemt=itemtypes(speed))
r3=markovdata(dat=speed[303:439,],itemt=itemtypes(speed))
summary(r2)
@

Here is the full specification of the \code{markovdata} function.

\input{markovdata}

\section{Data set speed}

Throughout this manual we will use a data set called speed, and hence
we provide some background information on how these data were
gathered.

\input{speed}

Interesting hypotheses to test are: is the switching regime symmetric?
Is there evidence for two states or does one state suffice?  Is the
guessing state actually a guessing state, i.e., is the probability
correct at chance level of 0.5?

\section{Defining models}

A dependent mixture model is defined by the number of states, and by
the item distribution functions, and can be created with the
\code{dmm}-function as follows:

<<>>= 
mod <-dmm(nstates=2,itemtypes=c("gaus",2))
summary(mod)
@

Here \code{itemtypes} is a vector of length the number of items
measured at each occasion specifying the desired distributions, in
this case the first item is to follow a normal distribution, and the
second item follows a bernouilli distribution.  Allowable
distributions are listed in Table~\ref{tbl:dist}, along with their
internal code, and the parameter names.  The R-internal code is used
for estimating these parameters.  Specifics of these distributions and
their estimation can be found in their respective help files.
Itemtypes can be specified by their name or by their internal code,
except in the case of multinomial items, which have to be specified by
a number.

\begin{table}[tbp]
	\centering
	\begin{tabular}{lcc}
		\hline
		distribution & code & parameters \\
		\hline
		\hline
		multinomial & $2,3,4, \ldots $ & $p_1, p_2, p_3, \ldots $  \\
		\hline
		gaussian, normal & 1  & $ \mu, \sigma $ \\
		\hline
		lognormal & -21 & $l\mu$ , $l\sigma$  \\
		\hline
		weibull & -22 & shape (a), scale (b) \\
		\hline
		gamma & -23 & shape (a), scale (s)  \\
		\hline
		3lognormal & -31 &  $l\mu$ , $l\sigma$, shift \\
		\hline
		3weibull & -32 & shape (a), scale (b), shift \\
		\hline
		3gamma & -33 & shape (a), scale (s), shift \\
		\hline
	\end{tabular}
	\caption{Allowable distribution names, internal codes, and number 
	of parameters.}
	\label{tbl:dist}
\end{table}

The function \code{dmm} returns an object of class \code{dmm} which
has its own summary function providing the parameter values of the
model. See the help files for further details. Except in simple 
cases, starting values can be a problem in latent Markov models, and
so in general it's best to provide them if you have a fairly good idea
of what to expect. Providing starting values is done through the stval
argument:

<<>>=
st <- c(1,0.9,0.1,0.2,0.8,2,1,0.7,0.3,5,2,0.2,0.8,0.5,0.5)
mod <- dmm(nsta=2,itemt=c(1,2), stval=st)
summary(mod)
@

\input{dmm}

\subsection{Generating data}

The \code{dmm}-class has a \code{generate} method that can be used to 
generate data according to a specified model.

<<>>=
gen<-generate(c(100,50),mod)
summary(gen)
@
\begin{figure}[htbp]
  \begin{center}
<<fig=TRUE>>=
plot(gen)
@
	\caption{Two timeseries generated by 2-state model with one 
	gaussian item and one binary item.}
  \end{center}
\end{figure}

\input{generate}

\section{Fitting models\label{fitmod}}

Fitting models is done using the function \code{fitdmm}. The standard
call only requires a dataset and a model as in:

<<>>= 
data(speed)
mod <- dmm(nstates=2,itemtypes=c(1,2))
fitex <- fitdmm(speed,mod)
@

Calling \code{fitdmm} produces some online ouput about the progress of
the optimization which can be controlled with the \code{printlevel}
argument.  Its default value of 1 just gives the first and the last
iteration of the optimization; 0 gives no output, and setting it to 5
or higher values will produce output at each iteration.  These values
correspond with the 0,1, and 2 printlevel of \code{nlm}.  When using
\code{optim}, the \code{printlevel} argument is used to set the
\code{REPORT} argument of \code{optim} (see its help page for
details).  Printlevel 0 gives report 0, printlevel 1 gives report 10,
printlevels 2--4 give report 5 and printlevel>4 gives report 1,
producing output at every iteration.  Printlevels starting from 15 and
higher produce increasingly annoying output from the C-routines that
compute the loglikelihood.

\code{Fitdmm} returns an object of class \code{fit} which has a
summary method showing the estimated parameter values, along with
standard errors, and t-ratios whenever those are available.  Along
with the log-likelihood, the AIC and BIC values are provided.  Apart
from the printed values (see summary below), a \code{fit}-object has a
number of other fields.  Most importantly, it contains a copy of the
fitted model in \code{mod} and it has a field \code{post} containing
posterior state estimates.  That is, for each group $g$,
\code{post\$states[[g]]} is a matrix with dimensions the number of
states of the model + 2, and \code{sum(ntimes(dat))}.  The first
column contains the a posteriori component model, the second column
has the state number within the component, and the other columns are
used for the a posteriori probabilities of each of the states.

<<>>=
summary(fitex)
@

\input{fitdmm}

\chapter{Extending and constraining models}

\section{Fixing and constraining parameters}

Continuing the example from above, it can be seen that in one of the
states, the probability of a correct answer is about .5, as is the
probability of an incorrect answer, i.e., these are parameters Item2,p1 and
Item2,p2. This latent state, is supposed to be a guessing state, and hence 
it makes sense to constrain these parameters to their theoretical values of 
.5. Similarly, the initial state probability for the slow state is one, and
zero for the other state, and hence it makes sense to fix these parameters.
The third constraint that we consider here is an equality constraint
between the transition parameters. Using this constraint, we can test the
hypothesis whether the switching between states is a symmetric process or
not. Hence, we constrain the transition parameters $a_{11}$ and $a_{22}$.

Constraining and fixing parameters is done in a similar fashion as the
\code{pa} command that is used in LISREL \citep{Lis99}. The
\code{conpat} argument to the \code{fitdmm}-function specifies for
each parameter in the model whether it's fixed (0) or free (1 or
higher). Equality constraints can be imposed by having two parameters have 
the same number in the \code{conpat} vector. When only fixed values are
required the \code{fixed} argument can be used instead of \code{conpat},
with zeroes for fixed parameters and other values (ones e.g.) for non-fixed
parameters.

Fitting the models subject to these constraints is mostly done through
reparametrization.  Inequality constraints are enforced by adding a
penalty to the loglikelihood when the constraint is not satisfied.
The penalty is linear in the amount by which the constraint is not
satisfied, and not logarithmic or something similar which is often
used (see e.g.\ the documentation for \code{constrOptim} which uses a
logarithmic boundary for inequality constraints).  This has advantages
and disadvantages.  There are two marked disadvantages.  First, the
loglikelihood is not smooth at the boundary of the paramter space.
Second, it can happen that the constraint is not satisfied.  Whenever
cosntraints are not satisfied \code{fitdmm} exits with a warning
stating the amount by which it is not satisfied.  This can be remedied
by upping the \code{vfactor} argument which simply increases the
penalty by this factor (its default value is 5).  An advantage is that
using a linear penalty, it is possible that the parameter is estimated
at the boundary, which is prohibited with logarithmic boundaries.

\paragraph{Parameter numbering} When using the \code{conpat} and
\code{fixed} arguments, complete vectors should be supplied, i.e., these
vectors should have length of the number of parameters of the model.
Parameters are numbered in the following order:
\begin{enumerate}
	\item  the mixing proportions of a mixture of latent Markov
	models, i.e., just one parameter for a single component model
	which has value 1 and is fixed

	\item  the component parameters for each component consisting of
	the following:
	\begin{enumerate}
		\item  transition parameters in row major order, $a_{11},
		a_{12}, a_{13}, \ldots, a_{21}, a_{22}, a_{23}, \ldots $

		\item  the observation parameters per state and per item, in
		the order listed in Table~\ref{tbl:dist}

		\item  the initial state probabilities per state
	\end{enumerate}

\end{enumerate}

<<>>=
conpat=rep(1,15)
conpat[1]=0 
conpat[14:15]=0 
conpat[8:9]=0
conpat[2]=conpat[5]=2
stv=c(1,.896,.104,.084,.916,5.52,.20,.5,.5,6.39,.24,.098,.902,0,1)
mod=dmm(nstates=2,itemt=c("n",2),stval=stv,conpat=conpat)
@

In the example above \code{conpat} is used to specify a number of
constraints.  First, \code{conpat[1]=0} specifies that the mixing
proportion of the model should be fixed (at its starting value of 1),
which is always the case for single component models.  Second,
\code{conpat[14:15]=0} fixes the initial state probabilities to zero
and one respectively.  Similarly, for \code{conpat[8:9]=0}, which are
the guessing state parameters for the accuracy scores.  They are both
fixed at 0.5 so as to make the guessing state an actual guessing
state.  Finally, by invoking \code{conpat[2]=conpat[5]=2}, transition
parameters $a_{11}$ and $a_{22}$ are set to be equal.  Whenever
equality constraints are not sufficient, general linear constraints
can be specified using the \code{conrows} argument.

The constrained model has the following estimated
parameters\footnote{Note that in running this example with the
starting values from the unconstrained model, the initial
loglikelihood is worse than the final loglikelihood because the
initial likelihood is based on parameters that do not satisfy the
constraints.}:

% <<results=hide>>=
% fitfix <- fitdmm(dat=speed,dmm=mod)
% @
% results in models.Rda

<<>>=
summary(fitfix)
@

\section{Multi group/case analysis}

<<echo=FALSE>>=
conpat=rep(1,15)
conpat[1]=0
conpat[8:9]=0
conpat[14:15]=0
stv=c(1,0.9,0.1,0.1,0.9,5.5,0.2,0.5,0.5,6.4,0.25,0.9,0.1,0.5,0.5)
mod<-dmm(nstates=2,itemt=c("n",2),stval=stv,conpat=conpat)
@

\code{depmix} can handle multiple cases or multiple groups. A
multigroup model is specified using the function \code{mgdmm} as
follows:

<<>>=
mgr <- mgdmm(dmm=mod,ng=3,trans=TRUE,obser=FALSE)
mgrfree <- mgdmm(dmm=mod,ng=3,trans=FALSE)
@

The \code{ng} argument specifies the number of groups, and the
\code{dmm} argument specifies the model for each group.  \code{dmm}
can be either a single model or list of models of \code{length(ng)}.
If it is a single model, each group has an identical structural model
(same fixed and constrained parameters), and else each group has its
model. Three further arguments can be used to constrain parameters
between groups, \code{trans}, \code{obser}, and \code{init}
respectively. By setting either of these to \code{TRUE}, the
corresponding transition, observation, and initial state parameters
are estimated equal between groups\footnote{There is at this moment no way of 
fine-tuning this to restrict equalities to individual parameters.
However, this can be accomplished by manually changing the linear
constraint matrix, and the corresponding upper and lower boundaries.}.

In this example, the model from above was used and fitted on the three
observed series, and the \code{trans=TRUE} ensures that the
transition matrix parameters are constrained to be equal between the
models for these series, whereas the observation parameters are freely
estimated, i.e.\ to capture learning effects. The resulting
parameters are:

% <<>>=
% fitmg <- fitdmm(dat=list(r1,r2,r3),dmm=mgr,ses=FALSE)
% fitmgfree <- fitdmm(dat=list(r1,r2,r3),dmm=mgrfree,ses=FALSE)
% @

<<>>=
summary(fitmg)
@

The loglikelihood ratio statistic can be used to test whether
constraining these transition parameters significantly reduces the
goodness-of-fit of the model.  The statistic has value
$LR=\Sexpr{round(2*(fitmgfree$loglike-fitmg$loglike),3)}$, and it has
an approximate $\chi^{2}$ distribution with $df=4$ because in each but
the first model, two transition matrix parameters were estimated equal
to the parameters in the first model (note that the other two
transition parameters were already had to be constrained to ensure
that the rows of the transition matrices sum to 1).  The associated
$p$-value for the statistic is
$p=\Sexpr{round(pchisq(2*(fitmgfree$loglike-fitmg$loglike),4,lower.tail=FALSE),3)}$,
indicating that constraining the transition matrix parameters does not
significantly worsen the goodness-of-fit of the model.

\input{mgdmm}

\section{Models with time-dependent covariates}

Specifying a model with covariates is done by including two arguments 
in a call to \code{dmm} called \code{tdfix} and \code{tdst}, where
\code{td} means time dependent. \code{tdfix} is a logical vector of
length the number of parameters of the model, specifying which
parameters are to be estimated time-dependent. For an arbitrary
parameter $\lambda$, the model that is estimated has the form:
\begin{equation}
	\lambda_{t}=\lambda_{0}+\beta x_{t},
\end{equation}
where $\lambda_{0}$ is the intercept of the parameter, $\beta$ is the 
regression coefficient, and $x_{t}$ is the time-dependent covariate.
The covariate has to be scaled to lie between 0 and 1; this is
neccessary to be able to impose the right constraints on $\beta$ in
order to ensure that $\lambda_{t}$ is always appropriate, ie within
its lower and upper bounds (mostly 0 and 1 for multinomial item
parameters and transition parameters etc). The current version of
\code{depmix} does not have non-time-dependent covariates, which can
simply be faked by having $x_{t}$ be constant, and there is only
support for a single covariate.

In the example below, the transition parameters (numbers 2--5) are
defined to depend on the covariate which is the pay-off for accuracy. 
Providing starting values for the covariates is optional. If not
provided they are chosen at random around 0 which usually works just
fine.

<<>>=
conpat=rep(1,15)
conpat[1]=0
conpat[8:9]=0
conpat[14:15]=0
conpat[2]=2
conpat[5]=2
stv=c(1,0.9,0.1,0.1,0.9,5.5,0.2,0.5,0.5,6.4,0.25,0.9,0.1,0,1)
tdfix=rep(0,15)
tdfix[2:5]=1
tdst=rep(0,15)
tdst[2:5]=c(-0.4,0.4,0.15,-0.15)

mod<-dmm(nstates=2,itemt=c("n",2),stval=stv,conpat=conpat,tdfix=tdfix,tdst=tdst,modname="twoboth+cov")
@

% <<results=hide>>=
% fittd <-
% fitdmm(dat=speed,dmm=mod,tdcov=1,print=5,der=0,ses=0,vf=130)
% @
% results in models.Rda

<<>>=
summary(fittd)
@

\chapter{Special topics}

\section{Starting values}

Although providing your own starting values is preferable,
\pkg{depmix} has a routine for generating starting values using the
\code{kmeans}-function from the \pkg{stats}-package.  This will
usually provide reasonable starting values, but can be way off in a
number of cases.  First, for univariate categorical time series,
\code{kmeans} does not work at all, and \pkg{depmix} will provide a
warning.  Second, for multivariate series with unordered categorical
items with more than 2 categories, \code{kmeans} may provide good
starting values, but they may similarly be completely off, due to the
implicit assumption in \code{kmeans} that the categories are
indicating an underlying continuum.  Starting values using
\code{kmeans} are automatically provided when a model is specified
without starting values.  The argument \code{kmst} to the
\code{fitdmm}-function can be used to control this behavior.

Starting values of the paramaters, either user provided or generated,
are further boosted by using posterior estimates.  That is, first the
a posteriori latent states are generated from the current parameter
values for the data at hand.  Next, from the a posteriori latent
states, new parameter estimates are derived.  This is done by default
and can be controlled by the \code{postst} argument.  Provided that
the starting values were close to their true values, using this
procedure further pushes those parameters in the right direction.  If
however the original values were bad, this procedure may result in bad
estimates, i.e., optimization will lead to some non-optimal local
maximum of the loglikelihood.

\section{Finite mixtures and latent class models}

The function \code{lca} can be used to specify latent class models
and/or finite mixture models. It is simply a wrapper for the \code{dmm}
function, and all it does is adding appropriate numbers of zeroes and 
ones to the parameter specification vectors for starting values, fixed
values and linear constraints. When a model has class \code{lca} the
summary function does not print the transition matrix (because it is
fixed and/or not estimated).

\section{Mixtures of latent Markov models}

\code{depmix} provides support for fitting mixtures of latent Markov
models using the \code{mixdmm} function; it takes a list of
\code{dmm}'s as argument, possibly together with the starting values
for the mixing proportions for each component model. There's an
example in the helpfiles.

\input{mixdmm}

An example of fitting a mixture of \code{dmm}'s is in the \code{fitdmm}
helpfile. It fits the model in the example to data from a
discrimination learning experiment which is provided as data set
\code{discrimination}.

\input{discrimination}

\nocite{Raijmakers2001}
\nocite{Maas2005b}
\nocite{Schmittmann2005b}
\nocite{McCutcheon1987}

\bibliography{all}

\end{document}