### 2010-10-18 papers/jss/depmixS4.Rnw
Minor typo's corrected.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@440 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article,nojss]{jss}
\usepackage{amsmath,thumbpdf}
\graphicspath{{Figures/}}

\newcommand{\vc}{\mathbf}
\newcommand{\mat}{\mathbf}
\newcommand{\greekv}[1]{\mbox{\boldmath$\mathrm{#1}$}}
\newcommand{\greekm}[1]{\mbox{\boldmath$#1$}}

%% \usepackage{Sweave}
\SweaveOpts{engine = R, echo = TRUE, keep.source = TRUE, eps = FALSE}

%\VignetteIndexEntry{depmixS4: An R Package for Hidden Markov Models}
%\VignetteDepends{depmixS4,gamlss,gamlss.data}
%\VignetteKeywords{hidden Markov model, dependent mixture model, mixture model, constraints}
%\VignettePackage{depmixS4}

\author{Ingmar Visser\\University of Amsterdam \And 
        Maarten Speekenbrink\\University College London}
\Plainauthor{Ingmar Visser, Maarten Speekenbrink}

\title{\pkg{depmixS4}: An \proglang{R} Package for Hidden Markov Models}
\Plaintitle{depmixS4: An R Package for Hidden Markov Models}

\Abstract{	
  This introduction to the \proglang{R} package \pkg{depmixS4} is a (slightly)
  modified version of \cite{Visser+Speekenbrink:2010}, published in the
  \emph{Journal of Statistical Software}.

  \pkg{depmixS4} implements a general framework for defining and
  estimating dependent mixture models in the \proglang{R}
  programming language.  This includes standard Markov
  models, latent/hidden Markov models, and latent class and finite
  mixture distribution models.  The models can be fitted on mixed
  multivariate data with distributions from the \code{glm} family,
  the (logistic) multinomial, or the multivariate normal
  distribution.  Other distributions can be added easily, and an
  example is provided with the {\em exgaus} distribution.
  Parameters are estimated by the expectation-maximization (EM) algorithm or, when (linear)
  constraints are imposed on the parameters, by direct numerical
  optimization with the \pkg{Rsolnp} or \pkg{Rdonlp2} routines.   
}

\Keywords{hidden Markov model, dependent mixture model, mixture model, constraints}

\Volume{36}
\Issue{7}
\Month{August}
\Year{2010}
\Submitdate{2009-08-19}
\Acceptdate{2010-06-21}

\Address{
  Ingmar Visser\\
  Department of Psychology\\
  University of Amsterdam\\
  Roetersstraat 15\\
  1018 WB, Amsterdam, The Netherlands\\
  E-mail: \email{i.visser@uva.nl} \\
  URL: \url{http://www.ingmar.org/}
}

\begin{document}

<<echo=FALSE,results=hide>>=	
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE, digits = 4)
library("depmixS4")
@

\section{Introduction}

Markov and latent Markov models are frequently used in the social
sciences, in different areas and applications.  In psychology, they
are used for modelling learning processes; see \citet{Wickens1982},
for an overview, and e.g., \citet{Schmittmann2006}, for a recent
application.  In economics, latent Markov models are so-called regime
switching models (see e.g., \citealp{Kim1994} and
\citealp{Ghysels1994}).  Further applications include speech
recognition \citep{Rabiner1989}, EEG analysis \citep{Rainer2000}, and
genetics \citep{Krogh1998}.  In these latter areas of application,
latent Markov models are usually referred to as hidden Markov models.
See for example \citet{Fruhwirth2006} for an overview of hidden Markov
models with extensions.  Further examples of applications can be found
in e.g., \citet[][Chapter~1]{Cappe2005}.  A more gentle introduction
into hidden Markov models with applications is the book by
\citet{Zucchini2009}.

The \pkg{depmixS4} package was motivated by the fact that while Markov
models are used commonly in the social sciences, no comprehensive
package was available for fitting such models.  Existing software for
estimating Markovian models include \pkg{Panmark} \citep{Pol1996}, and for
latent class models \pkg{Latent Gold} \citep{Vermunt2003}.  These programs
lack a number of important features, besides not being freely
available.  There are currently some packages in \proglang{R} that
handle hidden Markov models but they lack a number of features that we
needed in our research.  In particular, \pkg{depmixS4} was designed to
meet the following goals:

\begin{enumerate}	
	\item to be able to estimate parameters subject to general
	linear (in)equality constraints;

	\item to be able to fit transition models with covariates, i.e.,
	to have time-dependent transition matrices;

	\item to be able to include covariates in the prior or initial
	state probabilities;

	\item to be easily extensible, in particular, to allow users to
	easily add new uni- or multivariate response distributions and
	new transition models, e.g., continuous time observation models.

\end{enumerate}

Although \pkg{depmixS4} was designed to deal with longitudinal or time
series data, for say $T>100$, it can also handle the limit case when
$T=1$.  In this case, there are no time dependencies between observed
data and the model reduces to a finite mixture or latent class model.
While there are specialized packages to deal with mixture data, as far
as we know these don't allow the inclusion of covariates on the prior
probabilities of class membership.  The possibility to estimate the
effects of covariates on prior and transition probabilities is a
distinguishing feature of \pkg{depmixS4}.
In Section~\ref{sec:2}, we provide an outline of the model and likelihood equations.

The \pkg{depmixS4} package is implemented using the \proglang{R} system
for statistical computing \citep{R2010} and is available from the
Comprehensive \proglang{R} Archive Network at \url{http://CRAN.R-project.org/package=depmixS4}.

\section{The dependent mixture model} \label{sec:2}

The data considered here have the general form $\vc{O}_{1:T}=
(O_{1}^{1}, \ldots, O_{1}^{m}$, $O_{2}^{1}, \ldots, O_{2}^{m}$,
\ldots, $O_{T}^{1}, \ldots, O_{T}^{m})$ for an $m$-variate time series
of length $T$.  In the following, we use $\vc{O}_{t}$ as shorthand for
$O_{t}^{1}, \ldots, O_{t}^{m}$.  As an example, consider a time series
of responses generated by a single participant in a psychological
response time experiment.  The data consists of three variables:
response time, response accuracy, and a covariate which is a pay-off
variable reflecting the relative reward for speeded and/or accurate
responding.  These variables are measured on 168, 134 and 137
occasions respectively (the first series of 168 trials is plotted in
Figure~\ref{fig:speed}).  These data are more fully described in
\citet{Dutilh2010}, and in the next section a number of example models
for these data is described.

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[t!]
\begin{center}
<<fig=TRUE,echo=FALSE,height=5,width=7,eps=FALSE>>=	
data("speed")
plot(as.ts(speed[1:168,]), main = "Speed-accuracy trade-off")
@
	\caption{Response times (rt), accuracy (corr) and pay-off 
	values (Pacc) for the first series of responses in dataset 
	\code{speed}.}
	\label{fig:speed}
\end{center}
\end{figure}

The latent Markov model is usually associated with data of this type,
in particular for multinomially distributed responses.  However,
commonly employed estimation procedures \citep[e.g.,][]{Pol1996}, are
not suitable for long time series due to underflow problems.  In
contrast, the hidden Markov model is typically only used for `long'
univariate time series \citep[][Chapter~1]{Cappe2005}.  We use the
term ``dependent mixture model" because one of the authors (Ingmar
Visser) thought it was time for a new name to relate these
models\footnote{Only later he found out that \citet{Leroux1992}
already coined the term dependent mixture models in an application
with hidden Markov mixtures of Poisson count data.}.

The fundamental assumption of a dependent mixture model is that at any
time point, the observations are distributed as a mixture with $n$
components (or states), and that time-dependencies between the
observations are due to time-dependencies between the mixture
components (i.e., transition probabilities between the components).
These latter dependencies are assumed to follow a first-order Markov
process.  In the models we are considering here, the mixture
distributions, the initial mixture probabilities and transition
probabilities can all depend on covariates $\vc{z}_t$.

In a dependent mixture model, the joint likelihood of observations
$\vc{O}_{1:T}$ and latent states $\vc{S}_{1:T} = (S_1,\ldots,S_T)$,
given model parameters $\greekv{\theta}$ and covariates $\vc{z}_{1:T}
= (\vc{z}_1,\ldots,\vc{z}_T)$, can be written as:
\begin{equation}
	\Prob(\vc{O}_{1:T},\vc{S}_{1:T}|\greekv{\theta},\vc{z}_{1:T}) =  
	\pi_{i}(\vc{z}_1) \vc{b}_{S_t}(\vc{O}_1|\vc{z}_{1})
	\prod_{t=1}^{T-1} a_{ij}(\vc{z}_{t}) \vc{b}_{S_t}(\vc{O}_{t+1}|\vc{z}_{t+1}),
\end{equation}
where we have the following elements:
\begin{enumerate}

	\item $S_{t}$ is an element of $\mathcal{S}=\{1\ldots n\}$, a set
	of $n$ latent classes or states.

	\item $\pi_{i}(\vc{z}_1) = \Prob(S_1 = i|\vc{z}_1)$, giving the
	probability of class/state $i$ at time $t=1$ with covariate
	$\vc{z}_1$.

	\item $a_{ij}(\vc{z}_t) = \Prob(S_{t+1}=j|S_{t}=i,\vc{z}_t)$,
	provides the probability of a transition from state $i$ to state
	$j$ with covariate $\vc{z}_t$.

	\item $\vc{b}_{S_t}$ is a vector of observation densities
	$b_{j}^k(\vc{z}_t) = \Prob(O_{t}^k|S_t = j, \vc{z}_t)$ that
	provide the conditional densities of observations $O_{t}^k$
	associated with latent class/state $j$ and covariate $\vc{z}_t$,
	$j=1, \ldots, n$, $k=1, \ldots, m$.

\end{enumerate}

For the example data above, $b_j^k$ could be a Gaussian distribution
function for the response time variable, and a Bernoulli distribution
for the accuracy variable.  In the models considered here,
both the transition probability functions $a_{ij}$ and the initial
state probability functions $\greekv{\pi}$ may depend on covariates as
well as the response distributions $b_{j}^{k}$.

\subsection{Likelihood}

To obtain maximum likelihood estimates of the model parameters, we
need the marginal likelihood of the observations.  For hidden Markov
models, this marginal (log-)likelihood is usually computed by the
so-called forward-backward algorithm \citep{Baum1966,Rabiner1989}, or
rather by the forward part of this algorithm.  \cite{Lystig2002}
changed the forward algorithm in such a way as to allow computing the
gradients of the log-likelihood at the same time.  They start by
rewriting the likelihood as follows (for ease of exposition the
dependence on the model parameters and covariates is dropped here):
\begin{equation}
	L_{T} = \Prob(\vc{O}_{1:T}) = \prod_{t=1}^{T} 
	\Prob(\vc{O}_{t}|\vc{O}_{1:(t-1)}), 
	\label{condLike}
\end{equation}
where $\Prob(\vc{O}_{1}|\vc{O}_{0}):=\Prob(\vc{O}_{1})$. Note that for a 
simple, i.e., observed, Markov chain these probabilities reduce to 
$\Prob(\vc{O}_{t}|\vc{O}_{1},\ldots, 
\vc{O}_{t-1})=\Prob(\vc{O}_{t}|\vc{O}_{t-1})$.
The log-likelihood can now be expressed as:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log[\Prob(\vc{O}_{t}|\vc{O}_{1:(t-1)}].
	\label{eq:condLogl}
\end{equation}

To compute the log-likelihood, \cite{Lystig2002} define the following 
(forward) recursion:
\begin{align}
	\phi_{1}(j) &:= \Prob(\vc{O}_{1}, S_{1}=j) = \pi_{j} \vc{b}_{j}(\vc{O}_{1}) 
	\label{eq:fwd1} \\
	\phi_{t}(j) &:= \Prob(\vc{O}_{t}, S_{t}=j|\vc{O}_{1:(t-1)}) %\\
	= \sum_{i=1}^{N} [\phi_{t-1}(i)a_{ij} \vc{b}_{j}(\vc{O}_{t})] \times 
(\Phi_{t-1})^{-1},
	\label{eq:fwdt} 
\end{align}
where $\Phi_{t}=\sum_{i=1}^{N} \phi_{t}(i)$. Combining 
$\Phi_{t}=\Prob(\vc{O}_{t}|\vc{O}_{1:(t-1)})$, and 
equation~(\ref{eq:condLogl}) gives the following expression for the 
log-likelihood:
\begin{equation}
	l_{T} = \sum_{t=1}^{T} \log \Phi_{t}.
	\label{eq:logl}
\end{equation}

\subsection{Parameter estimation}

Parameters are estimated in \pkg{depmixS4} using the expectation-maximization (EM) algorithm or
through the use of a general Newton-Raphson optimizer.  In the EM
algorithm, parameters are estimated by iteratively maximizing the
expected joint log-likelihood of the parameters given the observations and
states.  Let $\greekv{\theta} = (\greekv{\theta}_1,
\greekv{\theta}_2,\greekv{\theta}_3)$ be the general parameter vector
consisting of three subvectors with parameters for the prior model,
transition model, and response models respectively.  The joint
log-likelihood can be written as:
\begin{multline}
\log \Prob(\vc{O}_{1:T}, \vc{S}_{1:T}|\vc{z}_{1:T},\greekv{\theta}) = \log 
\Prob(S_1|\vc{z}_{1},\greekv{\theta}_1) 
+ \sum_{t=2}^{T} \log \Prob(S_t|S_{t-1},\vc{z}_{t-1},\greekv{\theta}_2) \\
+ \sum_{t=1}^{T} \log \Prob(\vc{O}_t|S_t,\vc{z}_t,\greekv{\theta}_3)
\end{multline}
This likelihood depends on the unobserved states $\vc{S}_{1:T}$. In the 
Expectation step, we replace these with their expected values given a set of 
(initial) parameters $\greekv{\theta}' = (\greekv{\theta}'_1, 
\greekv{\theta}'_2,\greekv{\theta}'_3)$ and observations $\vc{O}_{1:T}$. 
The expected log-likelihood:
\begin{equation}
Q(\greekv{\theta},\greekv{\theta}') = E_{\greekv{\theta}'} 
(\log \Prob(\vc{O}_{1:T},\vc{S}_{1:T}|\vc{O}_{1:T},\vc{z}_{1:T},\greekv{\theta})),
\end{equation}
can be written as:
\begin{multline}
\label{eq:Q}
Q(\greekv{\theta},\greekv{\theta}') = 
\sum_{j=1}^n \gamma_1(j) \log \Prob(S_1=j|\vc{z}_1,\greekv{\theta}_1) \\ 
+ \sum_{t=2}^T \sum_{j=1}^n \sum_{k=1}^n \xi_t(j,k) \log \Prob(S_t = k|S_{t-1} 
= j,\vc{z}_{t-1},\greekv{\theta}_2)  \\
 + \sum_{t=1}^T \sum_{j=1}^n \sum_{k=1}^m \gamma_t(j) 
\log \Prob(O^k_t|S_t=j,\vc{z}_t,\greekv{\theta}_3),
\end{multline}
where the expected values $\xi_t(j,k) = P(S_t = k, S_{t-1} =
j|\vc{O}_{1:T}, \vc{z}_{1:T},\greekv{\theta}')$ and $\gamma_t(j) =
P(S_t = j|\vc{O}_{1:T}, \vc{z}_{1:T},\greekv{\theta}')$ can be
computed effectively by the forward-backward algorithm \citep[see
e.g.,][]{Rabiner1989}.  The Maximization step consists of the
maximization of (\ref{eq:Q}) for $\greekv{\theta}$.  As the right hand
side of (\ref{eq:Q}) consists of three separate parts, we can maximize
separately for $\greekv{\theta}_1$, $\greekv{\theta}_2$ and
$\greekv{\theta}_3$.  In common models, maximization for
$\greekv{\theta}_1$ and $\greekv{\theta}_2$ is performed by the
\code{nnet.default} routine in the \pkg{nnet} package
\citep{Venables2002}, and maximization for $\greekv{\theta}_3$ by the
standard \code{glm} routine.  Note that for the latter maximization,
the expected values $\gamma_t(j)$ are used as prior weights of the
observations $O^k_t$.

The EM algorithm however has some drawbacks.  First, it can be slow to
converge towards the end of optimization.  Second, applying
constraints to parameters can be problematic; in particular, EM can
lead to wrong parameter estimates when applying constraints.  Hence,
in \pkg{depmixS4}, EM is used by default in unconstrained models, but
otherwise, direct optimization is used.  Two options are available for
direct optimization using package \pkg{Rsolnp}
\citep{Ghalanos2010,Ye1987}, or \pkg{Rdonlp2}
\citep{Tamura2009,Spellucci2002}.  Both packages can handle general
linear (in)equality constraints (and optionally also non-linear
constraints).

\section[Using depmixS4]{Using \pkg{depmixS4}}

Two steps are involved in using \pkg{depmixS4} which are illustrated
below with examples:
\begin{enumerate}

	\item model specification with function \code{depmix} (or with
	\code{mix} for latent class and finite mixture models, see example
	below on adding covariates to prior probabilities);

	\item  model fitting with function \code{fit}.
\end{enumerate}
We have separated the stages of model specification and model fitting
because fitting large models can be fairly time-consuming and it is
hence useful to be able to check the model specification before
actually fitting the model.

\subsection[Example data: speed]{Example data: \code{speed}}

Throughout this article a data set called \code{speed} is used.  As
already indicated in the introduction, it consists of three time
series with three variables: response time \code{rt}, accuracy
\code{corr}, and a covariate, \code{Pacc}, which defines the relative
pay-off for speeded versus accurate responding.  Before describing
some of the models that are fitted to these data, we provide a brief
sketch of the reasons for gathering these data in the first place.

Response times are a very common dependent variable in psychological
experiments and hence form the basis for inference about many
psychological processes.  A potential threat to such inference based
on response times is formed by the speed-accuracy trade-off: different
participants in an experiment may respond differently to typical
instructions to `respond as fast and accurate as possible'.  A popular
model which takes the speed-accuracy trade-off into account is the
diffusion model \citep{Ratcliff1978}, which has proven to provide
accurate descriptions of response times in many different settings.

One potential problem with the diffusion model is that it predicts a
continuous trade-off between speed and accuracy of responding, i.e.,
when participants are pressed to respond faster and faster, the
diffusion model predicts that this would lead to a gradual decrease in
accuracy.  The \code{speed} data set that we analyze below was
gathered to test this hypothesis versus the alternative hypothesis
stating that there is a sudden transition from slow and accurate
responding to fast responding at chance level.  At each trial of the
experiment, the participant is shown the current setting of the
relative reward for speed versus accuracy.  The bottom panel of
Figure~\ref{fig:speed} shows the values of this variable.  The
experiment was designed to investigate what would happen when this
reward variable changes from reward for accuracy only to reward for
speed only.  The \code{speed} data that we analyse here are from
participant A in Experiment 1 in \citet{Dutilh2010}, who provide a
complete description of the experiment and the relevant theoretical
background.

The central question regarding this data is whether it is indeed best
described by two modes of responding rather than a single mode of
responding with a continuous trade-off between speed and accuracy.
The hallmark of a discontinuity between slow versus speeded responding
is that switching between the two modes is asymmetric \citep[see
e.g.][for a theoretical underpinning of this claim]{Maas1992}.  The
\code{fit} help page of \pkg{depmixS4} provides a number of examples
in which the asymmetry of the switching process is tested; those
examples and other candidate models are discussed at length in
\citet{Visser2009b}.

\subsection{A simple model}

A dependent mixture model is defined by the number of states and the
initial state, state transition, and response distribution functions.
A dependent mixture model can be created with the
\code{depmix} function as follows:
<<>>=
library("depmixS4")
data("speed")
set.seed(1)
mod <- depmix(response = rt ~ 1, data = speed, nstates = 2,
  trstart = runif(4))
@

The first line of code loads the \pkg{depmixS4} package and
\code{data(speed)} loads the \code{speed} data set.  The line
\code{set.seed(1)} is necessary to get starting values that will
result in the right model, see more on starting values below.

The call to \code{depmix} specifies the model with a number of
arguments.  The \code{response} argument is used to specify the
response variable, possibly with covariates, in the familiar format
using formulae such as in \code{lm} or \code{glm} models.  The second
argument, \code{data = speed}, provides the \code{data.frame} in which
the variables from \code{response} are interpreted.  Third, the number
of states of the model is given by \code{nstates = 2}.

\paragraph{Starting values.} Note also that start values for the
transition parameters are provided in this call by the \code{trstart}
argument.  Starting values for parameters can be provided using three
arguments: \code{instart} for the parameters relating to the prior
probabilities, \code{trstart}, relating the transition models, and
\code{respstart} for the parameters of the response models.  Note that
the starting values for the initial and transition models as well as
multinomial logit response models are interpreted as {\em
probabilities}, and internally converted to multinomial logit
parameters (if a logit link function is used).  Start values can also
be generated randomly within the EM algorithm by generating random
uniform values for the values of $\gamma_{t}$ in (\ref{eq:Q}) at
iteration 0.  Automatic generation of starting values is not available
for constrained models (see below).

\paragraph{Fitting the model and printing results.} The \code{depmix}
function returns an object of class `\code{depmix}' which contains the
model specification, and not a fitted model.  Hence, the model needs
to be fitted by calling \code{fit}:
<<>>=
fm <- fit(mod)
@

The \code{fit} function returns an object of class
`\code{depmix.fitted}' which extends the `\code{depmix}' class, adding
convergence information (and information about constraints if these
were applied, see below).  The class has \code{print} and
\code{summary} methods to see the results.  The \code{print} method
provides information on convergence, the log-likelihood and
the AIC and BIC values:
<<>>=
fm 
@

These statistics can also be extracted using \code{logLik}, \code{AIC}
and \code{BIC}, respectively.  By comparison, a 1-state model for
these data, i.e., assuming there is no mixture, has a log-likelihood of
$-305.33$, and 614.66, and 622.83 for the AIC and BIC respectively.
Hence, the 2-state model fits the data much better than the 1-state
model.  Note that the 1-state model can be specified using
\code{mod <- depmix(rt ~ 1, data = speed, nstates = 1)}, although this model is
trivial as it will simply return the mean and standard deviation of
the \code{rt} variable.

The \code{summary} method of \code{fit}ted models provides the
parameter estimates, first for the prior probabilities model, second
for the transition models, and third for the response models.
<<>>=
summary(fm)
@

Since no further arguments were specified, the initial state, state
transition and response distributions were set to their defaults
(multinomial distributions for the first two, and a Gaussian
distribution for the response).  The resulting model indicates two
well-separated states, one with slow and the second with fast
responses.  The transition probabilities indicate rather stable
states, i.e., the probability of remaining in either of the states is
around 0.9.  The initial state probability estimates indicate that
state 1 is the starting state for the process, with a negligible
probability of starting in state 2.

\subsection{Covariates on transition parameters}

By default, the transition probabilities and the initial state
probabilities are parameterized using a multinomial model with an
identity link function.  Using a multinomial logistic model allows one
to include covariates on the initial state and transition
probabilities.  In this case, each row of the transition matrix is
parameterized by a baseline category logistic multinomial, meaning
that the parameter for the base category is fixed at zero
\citep[see][p.~267~ff., for multinomial logistic models and various
parameterizations]{Agresti2002}.  The default baseline category is the
first state.  Hence, for example, for a 3-state model, the initial
state probability model would have three parameters of which the first
is fixed at zero and the other two are freely estimated.
\citet{Chung2007} discuss a related latent transition model for
repeated measurement data ($T=2$) using logistic regression on the
transition parameters; they rely on Bayesian methods of estimation.
Covariates on the transition probabilities can be specified using a
one-sided formula as in the following example:
<<>>=
set.seed(1)
mod <- depmix(rt ~ 1, data = speed, nstates = 2, family = gaussian(),
  transition = ~ scale(Pacc), instart = runif(2)) 
fm <- fit(mod, verbose = FALSE) 
@

Note the use of \code{verbose = FALSE} to suppress printing of
information on the iterations of the fitting process.  Applying
\code{summary} to the \code{fit}ted object gives (only transition
models printed here by using argument \code{which}): 
<<>>= 
summary(fm, which = "transition") 
@

The summary provides all parameters of the model, also the (redundant)
zeroes for the base-line category in the multinomial model.  The
summary also prints the transition probabilities at the zero value of
the covariate.  Note that scaling of the covariate is useful in this
regard as it makes interpretation of these intercept probabilities
easier.

\subsection{Multivariate data}

Multivariate data can be modelled by providing a list of formulae as
well as a list of family objects for the distributions of the various
responses.  In above examples we have only used the response times
which were modelled as a Gaussian distribution.  The accuracy
variable in the \code{speed} data can be modelled with a multinomial
by specifying the following:
<<>>=
set.seed(1)
mod <- depmix(list(rt ~ 1,corr ~ 1), data = speed, nstates = 2, 
  family = list(gaussian(), multinomial("identity")),
  transition = ~ scale(Pacc), instart = runif(2))
fm <- fit(mod,verbose = FALSE)
@

This provides the following fitted model parameters (only the 
response parameters are given here): 
<<>>=
summary(fm, which = "response")
@

As can be seen, state 1 has fast response times and accuracy is
approximately at chance level (.474), whereas state 2 corresponds with
slower responding at higher accuracy levels (.904).

Note that by specifying multivariate observations in terms of a list,
the variables are considered conditionally independent (given the
states).  Conditionally \emph{dependent} variables must be handled as
a single element in the list.  Effectively, this means specifying a
multivariate response model.  Currently, \pkg{depmixS4} has one 
multivariate response model which is for multivariate normal
variables.

\subsection{Fixing and constraining parameters}

Using package \pkg{Rsolnp} \citep{Ghalanos2010} or \pkg{Rdonlp2}
\citep{Tamura2009}, parameters may be fitted subject to general linear
(in-)equality constraints.  Constraining and fixing parameters is done
using the \code{conpat} argument to the \code{fit} function, which
specifies for each parameter in the model whether it's fixed (0) or
free (1 or higher).  Equality constraints can be imposed by giving two
parameters the same number in the \code{conpat} vector.  When only
fixed values are required, the \code{fixed} argument can be used
instead of \code{conpat}, with zeroes for fixed parameters and other
values (e.g., ones) for non-fixed parameters.  Fitting the models
subject to these constraints is handled by the optimization routine
\code{solnp} or, optionally, by \code{donlp2}.  To be able to
construct the \code{conpat} and/or \code{fixed} vectors one needs the
correct ordering of parameters which is briefly discussed next before
proceeding with an example.

\paragraph{Parameter numbering.} When using the \code{conpat} and
\code{fixed} arguments, complete parameter vectors should be supplied,
i.e., these vectors should have length equal to the number of parameters of
the model, which can be obtained by calling \code{npar(object)}.  Note
that this is not the same as the degrees of freedom used e.g., in the
\code{logLik} function because \code{npar} also counts the baseline
category zeroes from the multinomial logistic models.  Parameters are
numbered in the following order:
\begin{enumerate}
	\item  the prior model parameters;
	\item  the parameters for the transition models;
	\item  the response model parameters per state (and subsequently
	per response in the case of multivariate time series).
\end{enumerate}

To see the ordering of parameters use the following:
<<term=FALSE>>=
setpars(mod, value = 1:npar(mod))
@

To see which parameters are fixed (by default only baseline parameters
in the multinomial logistic models for the transition models and the
initial state probabilities model):
<<term=FALSE>>=
setpars(mod, getpars(mod, which = "fixed"))
@
When fitting constraints it is useful to have good starting values 
for the parameters and hence we first fit the following model without
constraints:
<<>>=
trst <- c(0.9, 0.1, 0, 0, 0.1, 0.9, 0, 0)
mod <- depmix(list(rt ~ 1,corr ~ 1), data = speed, transition = ~ Pacc,
  nstates = 2, family = list(gaussian(), multinomial("identity")),
  trstart = trst, instart = c(0.99, 0.01))
fm1 <- fit(mod,verbose = FALSE)
@

After this, we use the fitted values from this model to constrain the
regression coefficients on the transition matrix (parameters number~6
and~10):
<<term=FALSE, echo=TRUE,results=hide>>=
pars <- c(unlist(getpars(fm1)))
pars[6] <- pars[10] <- 11
pars[1] <- 0
pars[2] <- 1
pars[13] <- pars[14] <- 0.5
fm1 <- setpars(mod, pars)
conpat <- c(0, 0, rep(c(0, 1), 4), 1, 1, 0, 0, 1, 1, 1, 1)
conpat[6] <- conpat[10] <- 2
fm2 <- fit(fm1, equal = conpat)
@

Using \code{summary} on the fitted model shows that the regression
coefficients are now estimated at the same value of 12.66.  Setting
elements 13 and 14 of \code{conpat} to zero resulted in fixing those
parameters at their starting values of 0.5.  This means that state 1
can now be interpreted as a 'guessing' state which is associated with
comparatively fast responses.  Similarly for elements 1 and 2,
resulting in fixed initial probabilities.  The function \code{llratio}
computes the likelihood ratio $\chi^2$-statistic and the associated
$p$-value with appropriate degrees of freedom for testing the
tenability of constraints \citep{Dannemann2007}.  Note that these
arguments (i.e., \code{conpat} and \code{conrows}) provide the
possibility for arbitrary constraints, also between, e.g., a
multinomial regression coefficient for the transition matrix and the
mean of a Gaussian response model.  Whether such constraints make
sense is hence the responsibility of the user.

\subsection{Adding covariates on the prior probabilities}

To illustrate the use of covariates on the prior probabilities we have
included another data set with \pkg{depmixS4}.  The \code{balance}
data consists of 4 binary items (correct-incorrect) on a balance scale
task \citep{Siegler1981}.  The data form a subset of the data
published in \citet{Jansen2002}.  Before specifying specifying a model
for these data, we briefly describe them.

The balance scale task is a famous task for testing cognitive
strategies developed by Jean Piaget \citep[see][]{Siegler1981}.
Figure~\ref{fig:balance} provides an example of a balance scale item.
Participants' task is to say to which side the balance will tip when
released, or alternatively, whether it will stay in balance.  The item
shown in Figure~\ref{fig:balance} is a so-called distance item: the
number of weights placed on each side is equal, and only the distance
of the weights to the fulcrum differs between each side.

\setkeys{Gin}{width=0.7\textwidth}
\begin{figure}[t!]
\begin{center}
	\includegraphics{baldist.pdf}
	\caption{Balance scale item; this is a distance item (see the text 
	for details).}
	\label{fig:balance}  
\end{center}
\end{figure}

Children in the lower grades of primary school are known to ignore the
distance dimension, and base their answer only on the number of
weights on each side.  Hence, they would typically provide the wrong
answer to these distance items.  Slightly older children do take
distance into account when responding to balance scale items, but they
only do so when the number of weights is equal on each side.  These
two strategies that children employ are known as Rule~I and Rule~II.
Other strategies can be teased apart by administering different items.
The \code{balance} data set that we analyse here consists of 4
distance items on a balance scale task administered to 779
participants ranging from 5 to 19 years of age.  The full set of items
consisted of 25 items; other items in the test are used to detect
other strategies that children and young adults employ in solving
balance scale items \citep[see][for details]{Jansen2002}.

In the following model, age is included as covariate on class
membership to test whether, with age, children apply more complex
rules in solving balance scale items.  Similarly to the transition
matrix, covariates on the prior probabilities of the latent states (or
classes in this case), are defined by using a one-sided formula
\code{prior = ~ age}: 
<<>>= 
data("balance")
set.seed(1)
mod <- mix(list(d1 ~ 1, d2 ~ 1, d3 ~ 1, d4 ~ 1), data = balance,
  nstates = 3, family = list(multinomial("identity"),
  multinomial("identity"), multinomial("identity"),
  multinomial("identity")), respstart = runif(24), prior = ~ age,
  initdata = balance)
fm <- fit(mod, verbose = FALSE) 
fm
@
Note here that we define a \code{mix} model instead of a \code{depmix}
model as these data form independent observations.  More formally,
\code{depmix} models extend the class of `\code{mix}' models by adding
transition models.  As for fitting \code{mix} models: as can be
seen in Equation~\ref{eq:Q}, the EM algorithm can be applied by simply
dropping the second summand containing the transition parameters, and 
this is implemented as such in the EM algorithms in \pkg{depmixS4}.

As can be seen from the print of the fitted model above, the BIC for 
this model equals 1941.6. The similarly defined 2-class model for 
these data has a BIC of 1969.2, and the 4-class model has BIC equal to
1950.4. Hence, the 3-class seems to be adequate for describing these 
data.

The summary of the \code{fit}ted model gives the following (only the
prior model is shown here):
<<>>=
summary(fm, which = "prior")
@

The intercept values of the multinomial logistic parameters provide
the prior probabilities of each class when the covariate has value
zero (note that in this case the value zero does not have much
significance, scaling and/or centering of covariates may be useful in
such cases).  The summary function prints these values.  As can be
seen from those values, at age zero, the prior probability is
overwhelmingly at the second class.  Inspection of the response
parameters reveals that class~2 is associated with incorrect
responding, whereas class~1 is associated with correct responding;
class~3 is an intermediate class with guessing behavior.
Figure~\ref{fig:balprior} depicts the prior class probabilities as
function of age based on the fitted parameters.

\begin{figure}[htbp]
\begin{center}
<<fig=TRUE,echo=FALSE,height=5,width=7,eps=FALSE>>=
x <- mlogit(base=1)
coeff <- coefficients(fm@prior@parameters)

pr1 <- function(y) {sapply(y, function(z) {x$linkinv(c(t(coeff)%*%c(1,z)), base=1)[1]})}
pr2 <- function(y) {sapply(y, function(z) {x$linkinv(c(t(coeff)%*%c(1,z)), base=1)[2]})}
pr3 <- function(y) {sapply(y, function(z) {x$linkinv(c(t(coeff)%*%c(1,z)), base=1)[3]})}

plot(pr1,min(balance$age),max(balance$age),lty=1,ylim=c(0,1),
main="Prior probabilities by age, balance scale data", xlab="age", ylab="Pr")
plot(pr2,min(balance$age),max(balance$age),add=T,lty=2)
plot(pr3,min(balance$age),max(balance$age),add=T,lty=3)

legend("right",legend=c("Class 1 (correct)","Class 2 (incorrect)","Class 3 (guess)"),lty=1:3,inset=c(0.1,0))
@
	\caption{Class probabilities as a function of age.}
	\label{fig:balprior}  
\end{center}
\end{figure}

As can be seen from Figure~\ref{fig:balprior}, at younger ages 
children predominantly apply Rule I, the wrong strategy for these 
items. According to the model, approximately 90 \% of children at age
5 apply Rule I. The remaining 10 \% are evenly split among the 2 
other classes. At age 19, almost all participants belong to class~1. 
Interestingly, prior probability of the 'guess' class 2, first 
increases with age, and then decreases again. This suggests that 
children pass through a phase in which they are uncertain or possibly 
switch between applying different strategies.

\section[Extending depmixS4]{Extending \pkg{depmixS4}}

The \pkg{depmixS4} package was designed with the aim of making it
relatively easy to add new response distributions (as well as possibly
new prior and transition models).  To make this possible, the EM
routine simply calls the \code{fit} methods of the separate response
models without needing access to the internal workings of these
routines.  Referring to equation~\ref{eq:Q}, the EM algorithm calls
separate fit functions for each part of the model, the prior
probability model, the transition models, and the response models.  As
a consequence, adding user-specified response models is
straightforward.

User-defined distributions should extend the `\code{response}' class and
have the following slots:
\begin{enumerate}
	\item \code{y}: The response variable.
	\item \code{x}: The design matrix, possibly only an intercept.
	\item \code{parameters}: A named list with the coefficients and possibly 
	other parameters (e.g., the standard deviation in the normal 
	response model).
	\item \code{fixed}: A vector of logicals indicating whether parameters are 
	fixed.
	\item \code{npar}: Numerical indicating the number of parameters of the model.
\end{enumerate}

In the \code{speed} data example, it may be more appropriate to model
the response times with an exgaus rather than a Gaussian distribution.
To do so, we first define an `\code{exgaus}' class extending the
`\code{response}' class:
<<>>=
setClass("exgaus", contains="response")
@

The so-defined class now needs a number of methods: 
\begin{enumerate}
	\item constructor: Function to create instances of the class 
	with starting values.
	\item \code{show}: To print the model to the terminal.
	\item \code{dens}: The function that computes the density of the responses.
	\item \code{getpars} and \code{setpars}: To get and set parameters .
	\item \code{predict}: To generate predicted values.
	\item \code{fit}: Function to fit the model using posterior weights (used 
	by the EM algorithm).
\end{enumerate}

Only the constructor and the \code{fit} methods are provided here; the
complete code can be found in the help file of the \code{makeDepmix}
function.  The example with the exgaus distribution uses the
\pkg{gamlss} and \pkg{gamlss.dist} packages \citep{Rigby2005,
Stasinopoulos2007,Stasinopoulos2009a,Stasinopoulos2009b} for computing
the \code{dens}ity and for \code{fit}ting the parameters.

The constructor method return an object of class `\code{exgaus}', and is
defined as follows:
<<results=hide, term=FALSE>>=
library("gamlss")
library("gamlss.dist")
setGeneric("exgaus", function(y, pstart = NULL, fixed = NULL, ...) 
  standardGeneric("exgaus"))

setMethod("exgaus", 
  signature(y = "ANY"), 
  function(y, pstart = NULL, fixed = NULL, ...) {
    y <- matrix(y, length(y))
    x <- matrix(1) 
    parameters <- list()
    npar <- 3
    if(is.null(fixed)) fixed <- as.logical(rep(0, npar))
    if(!is.null(pstart)) {
      if(length(pstart) != npar) stop("length of 'pstart' must be ", npar)
      parameters$mu <- pstart[1]
      parameters$sigma <- log(pstart[2])
      parameters$nu <- log(pstart[3])
    }
    mod <- new("exgaus", parameters = parameters, fixed = fixed,
      x = x, y = y, npar = npar)
    mod
  }
)
@

<<echo=FALSE, term=FALSE>>=
setMethod("dens","exgaus",
    function(object,log=FALSE) {
    	dexGAUS(object@y, mu = predict(object), 
	sigma = exp(object@parameters$sigma), 
	nu = exp(object@parameters$nu), 
	log = log)
    }
)

setMethod("getpars","response",
    function(object,which="pars",...) {
        switch(which,
            "pars" = {
                parameters <- numeric()
                parameters <- unlist(object@parameters)
                pars <- parameters
            },
            "fixed" = {
                pars <- object@fixed
            }
        )
        return(pars)
    }
)

setMethod("setpars","exgaus",
    function(object, values, which="pars", ...) {
        npar <- npar(object)
        if(length(values)!=npar) stop("length of 'values' must be",npar)
        # determine whether parameters or fixed constraints are being set
		nms <- names(object@parameters)
		switch(which,
		  "pars"= {
		      object@parameters$mu <- values[1]
		      object@parameters$sigma <- values[2]
		      object@parameters$nu <- values[3]
		      },
		  "fixed" = {
		      object@fixed <- as.logical(values)
		  }
	    )
        names(object@parameters) <- nms
        return(object)
    }
)

setMethod("predict","exgaus", 
    function(object) {
        ret <- object@parameters$mu
        return(ret)
    }
)
@

The \code{fit} method is defined as follows: 
<<>>=
setMethod("fit", "exgaus",
  function(object, w) {
    if(missing(w)) w <- NULL
    y <- object@y
    fit <- gamlss(y ~ 1, weights = w, family = exGAUS(),
      control = gamlss.control(n.cyc = 100, trace = FALSE),
      mu.start = object@parameters$mu,
      sigma.start = exp(object@parameters$sigma),
      nu.start = exp(object@parameters$nu))    
    pars <- c(fit$mu.coefficients, fit$sigma.coefficients, 
      fit$nu.coefficients)
    object <- setpars(object,pars)
    object
  }
)
@

The \code{fit} method defines a \code{gamlss} model with 
only an intercept to be estimated and then sets the fitted parameters 
back into their respective slots in the `exgaus' object. See the help 
for \code{gamlss.distr} for interpretation of these parameters.

After defining all the necessary methods for the new response model, 
we can  now define the dependent mixture model using this response model. 
The function \code{makeDepmix} is included in \pkg{depmixS4} to have 
full control over model specification, and we need it here.

We first create all the response models that we need as a double list: 
<<>>=
rModels <- list()
rModels[[1]] <- list()
rModels[[1]][[1]] <- exgaus(speed$rt, pstart = c(5, 0.1, 0.1))
rModels[[1]][[2]] <- GLMresponse(formula = corr ~ 1, data = speed,
  family = multinomial(), pstart = c(0.5, 0.5))

rModels[[2]] <- list()
rModels[[2]][[1]] <- exgaus(speed$rt, pstart = c(6, 0.1, 0.1))
rModels[[2]][[2]] <- GLMresponse(formula = corr ~ 1, data = speed, 
  family = multinomial(), pstart = c(0.1, 0.9))
@

Next, we define the transition and prior probability models using the 
\code{transInit} function (which produces a transInit model, which also extends 
the `\code{response}' class): 
<<>>=
trstart <- c(0.9, 0.1, 0.1, 0.9)
transition <- list()
transition[[1]] <- transInit(~ Pacc, nst = 2, data = speed, 
  pstart = c(0.9, 0.1, 0, 0))
transition[[2]] <- transInit(~ Pacc, nst = 2, data = speed, 
  pstart = c(0.1, 0.9, 0, 0))
inMod <- transInit(~ 1, ns = 2, pstart = c(0.1, 0.9),
  data = data.frame(1))
@

Finally, we put everything together using \code{makeDepmix} and fit 
the model: 
<<>>=
mod <- makeDepmix(response = rModels, transition = transition,
  prior = inMod, stat = FALSE)
fm <- fit(mod, verbose = FALSE)
@

Using \code{summary} will print the fitted parameters.  Note that the
use of \code{makeDepmix} allows the possibility of, say, fitting a
gaussian in one state and an exgaus distribution in another state.
Note also that according to the AIC and BIC, the model with the exgaus
describes the data much better than the same model in which the
response times are modelled as gaussian.

\section{Conclusions and future work}

\pkg{depmixS4} provides a flexible framework for fitting dependent
mixture models for a large variety of response distributions.  It can
also fit latent class regression and finite mixture models, although
it should be noted that more specialized packages are available for
this such as \pkg{flexmix} \citep{Leisch2004,GruenLeisch2008}.  The package is
intended for modelling (individual) time series data with the aim of
characterizing the transition processes underlying the data.  The
possibility to use covariates on the transition matrix greatly
enhances the flexibility of the model.  The EM algorithm uses a very
general interface that allows easy addition of new response models.

We are currently working on implementing the gradients for response
and transition models with two goals in mind.  First, to speed up
(constrained) parameter optimization using \pkg{Rdonlp2} or
\pkg{Rsolnp}.  Second, analytic gradients are useful in computing the
Hessian of the estimated parameters so as to arrive at standard errors
for those.  We are also planning to implement goodness-of-fit
statistics \citep{Titman2008}.

\section*{Acknowledgments}

Ingmar Visser was supported by an EC Framework 6 grant, project 516542
(NEST).  Maarten Speekenbrink was supported by ESRC grant
RES-062-23-1511 and the ESRC Centre for Economic Learning and Social
Evolution (ELSE).  Han van der Maas provided the speed-accuracy data
\citep{Dutilh2010} and thereby necessitated implementing models with
time-dependent covariates.  Brenda Jansen provided the balance scale
data set \citep{Jansen2002} which was the perfect opportunity to test
the covariates on the prior model parameters.  The examples in the
help files use both of these data sets.

\bibliography{depmixS4}

\end{document}